---
title: "Fisher Metagenomes - With Subscales - 12 Aug 2019 Data"
output: 
  html_document: 
    highlight: textmate
    keep_md: yes
---

## Setup

```{r setup-pkgs-scripts, include=FALSE}
source("packages_and_sources.R")
# source("../Aux_scripts/colorblind_pals.R")
packages.bibtext <- "flannery_stagaman_et_al_R_packages.bib"
write_bib((.packages()), file = packages.bibtext) # from knitr
# Set options
options(knitr.table.format = "html")
.pardefault <- par(no.readonly = T)
theme_set(theme_cowplot())
figure_theme <- theme_update(
  plot.background = element_blank(),
  plot.title = element_text(size = 12),
  plot.subtitle = element_text(face = "italic", size = 8),
  panel.background = element_blank(),
  panel.grid = element_blank(),
  axis.text = element_text(size = 6),
  axis.title = element_text(size = 10),
  strip.background = element_blank(),
  strip.text = element_text(size = 8),
  legend.title = element_text(size = 7),
  legend.text = element_text(size = 5),
  legend.box.background = element_blank(),
  legend.key = element_blank()
)
opts_chunk$set(
  echo = FALSE,
  tidy = TRUE,
  warning = FALSE,
  message = FALSE,
  # out.width = 650,
  # out.height = 650 * 0.6,
  dpi = 300
)

```

```{r setup-vars-data, include=FALSE}
###### SET VARIABLES ######
main.seed <- 1337
set.seed(main.seed)
nproc <- 3
pnas.1col <- 3.42 # inches
pnas.2col <- 7 # inches
pnas.max.ht <- 9 # inches
outer.pt.sz <- 3.8
inner.pt.sz <- outer.pt.sz / 2
lab.font.pt <- 6
lab.size <- lab.font.pt * 0.35
seg.size <- 0.75
cb.pal <- getCB.pal("gray")
blue_to_red <- c("#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
purple_to_green <- c("#7b3294", "#c2a5cf", "#a6dba0", "#008837")
dist.method.name <- "Bray-Curtis"
dist.method <- "bray"
ord.method <- "PCoA"
nAxes <- 3 # number of axes to generate in the PCoA ordinations
kegg.remap <- FALSE
kegg.map.files <- c(
  "Static_data/map_kos_mod.rds",
  "Static_data/my_kos_mods_clpsd.rds"
)
meta.data.file <- "Static_data/stool_combine_diet_729_2019-08-12.csv"
# check filetype of metadata file
if (grepl(".txt", meta.data.file)) {
  data.file.sep <- "\t"
} else if (grepl(".csv", meta.data.file)) {
  data.file.sep <- ","
} else {
  stop("Meta-data file type not recognized")
}
data.date <- {
  strsplit(meta.data.file, split = ".", fixed = T)[[1]][1] %>% strsplit(split = "_")
}[[1]] %>%
  tail(1) # grab date from metadata file name
# name and create directories for saving objects and plots
saveDir <- paste("Saved_objs", data.date, "data", sep = "_")
if (!dir.exists(saveDir)) {
  dir.create(saveDir)
  cat(paste(saveDir, "created"))
}
# figDir <- "Figure_plots"
# if (!dir.exists(figDir)) {
#   dir.create(figDir)
#   cat(paste(figDir, "created"))
# }

###### IMPORT DATA ######
merged.abund.stats <- read.csv("Static_data/merged_abund_stats.csv", header = TRUE) # abundance stats file
merged.rel.abunds <- readRDS("Static_data/shotmap_merged_rel_abunds.rds") # KO relative abundances data
metaphlan.tax.tbl <- readRDS("Static_data/metaphlan_lowest_tax_tbl.rds") # taxon abundance table
# file containing covar names and descriptions
vars.legend <- read.table(
  "Static_data/variables_legend_2019-08-12.txt",
  sep = "\t",
  header = TRUE,
  na.strings = "N/A",
  stringsAsFactors = FALSE
) %>% as.data.table()
setkey(vars.legend, var.name)
# a data.frame with covariate categories
varIDs <- data.table(
  "group" = c(
    "intr",           "intr",       "ctrl", "intr",        "ctrl",  "ctrl"
  ),
  "varID" = c(
    "parent_",   "child_behav_",      "demo_", "esa_",        "gut_", "Diet_"
  ),
  "name" =  c(
    "Parenting", "Child Behavior", "Demography",  "ESA", "Gut-related",  "Diet"
  ),
  stringsAsFactors = FALSE
)
# grab only stool samples
stool.kos.rel.abunds <- merged.rel.abunds[grepl("SP", row.names(merged.rel.abunds)), ]
# remove sample SP001 in favor of sample SP001_2 which had better coverage (sample sample source)
stool.kos.rel.abunds <- stool.kos.rel.abunds[-c(3, 4), ]
# rename rows to simple SP003_R1, SP003_R2 format
row.names(stool.kos.rel.abunds) <- paste(
  sapply(strsplit(row.names(stool.kos.rel.abunds), "_S"), `[`, 1),
  c(str_match(row.names(stool.kos.rel.abunds), "R[12]")),
  sep = "_"
)
# import metadata
stool.metadata <- read.table(
  meta.data.file,
  header = TRUE,
  row.names = 2,
  sep = data.file.sep,
  na.strings = "NA",
  stringsAsFactors = FALSE
)
stool.metadata <- stool.metadata[, -1]
dropCols <- grep(
  "_per_sample|grains*_to_|persample|_av_|probiotic", 
  names(stool.metadata), 
  value = T
)
stool.metadata <- stool.metadata[, -match(dropCols, names(stool.metadata))]

###### CREATE PHYLOSEQ OBJECTS ######
# create a phyloseq object of all reads and their KO assignments
(stool.kos.phy <- phyloseq(
  sample_data(stool.metadata),
  otu_table(as.matrix(stool.kos.rel.abunds), taxa_are_rows = FALSE)
))
# remove samples with fewer than 1,000,000 total reads
(stool.kos.phy <- subset_samples(stool.kos.phy, Total.Processed.Reads >= 1000000))

# create a phyloseq object of just the foward reads and their KO assignments
# remove KOs not present in any remaining samples
stool1.kos.phy <- subset_samples(stool.kos.phy, read == "R1")
(stool1.kos.phy <- prune_taxa(taxa_sums(stool1.kos.phy) > 0, stool1.kos.phy))
saveRDS(stool1.kos.phy, file = file.path(saveDir, "phyloseq_stool1_kos.rds"))

# create a phyloseq object of just the reverse reads and their KO assignments
# remove KOs not present in any remaining samples
stool2.kos.phy <- subset_samples(stool.kos.phy, read == "R2")
(stool2.kos.phy <- prune_taxa(taxa_sums(stool2.kos.phy) > 0, stool2.kos.phy))
saveRDS(stool2.kos.phy, file = file.path(saveDir, "phyloseq_stool2_kos.rds"))

# match remaining samples to the taxon table
stool.tax.tbl <- metaphlan.tax.tbl[grepl("SP", rownames(metaphlan.tax.tbl)), ]
# create a phyloseq object of all reads and their taxon assignments
stool.tax.phy <- phyloseq(
  sample_data(stool.metadata),
  otu_table(stool.tax.tbl, taxa_are_rows = FALSE)
)
# remove samples with fewer than 1,000,000 total reads
(stool.tax.phy <- subset_samples(stool.tax.phy, Total.Processed.Reads >= 1000000))

# create a phyloseq object of just the foward reads and their taxon assignments
# remove taxa not present in any remaining samples
stool1.tax.phy <- subset_samples(stool.tax.phy, read == "R1")
(stool1.tax.phy <- prune_taxa(taxa_sums(stool1.tax.phy) > 0, stool1.tax.phy))
# Split taxon assignments into a table with taxon levels (Phylum to Strain) as columns
taxa.names <- taxa_names(stool1.tax.phy)
tax.tbl <- do.call(
  "rbind",
  lapply(
    taxa.names,
    function(taxon) {
      tax.split <- strsplit(taxon, "|", fixed = T)[[1]]
      if (length(tax.split) < 8) {
        tax.split <- c(tax.split, rep(NA, (8 - length(tax.split))))
      }
      tax.row <- sub("\\w__", "", tax.split, perl = T)
    }
  )
)
colnames(tax.tbl) <- c(
  "Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species", "Strain"
)
rownames(tax.tbl) <- taxa.names
# View(tax.tbl)
tax_table(stool1.tax.phy) <- tax_table(tax.tbl)
saveRDS(stool1.tax.phy, file = file.path(saveDir, "phyloseq_stool_r1s_metagenome_taxa.rds"))

# create a phyloseq object of just the reverse reads and their taxon assignments
# remove taxa not present in any remaining samples
stool2.tax.phy <- subset_samples(stool.tax.phy, read == "R2")
(stool2.tax.phy <- prune_taxa(taxa_sums(stool2.tax.phy) > 0, stool2.tax.phy))
tax_table(stool2.tax.phy) <- tax_table(tax.tbl)
saveRDS(stool2.tax.phy, file = file.path(saveDir, "phyloseq_stool_r2s_metagenome_taxa.rds"))

# Remap KOs to Modules if required or desired
if (!all(file.exists(kegg.map.files)) | kegg.remap) {
  source("Aux_scripts/kegg_maps.R")
}
```

## 1. KOs and taxonomic PCoA ordinations

The four PCoA ordinations below represent represent the Bray-Curtis distances for the stool samples based on the forward read data (read1s, plots on left), the reverse read data (read2s; plots on right). The distances were computed for their functional diversity (`shotmap`; KEGG Orthologies; plots on top) and taxonomic diversity (`metaphlan2`; plots on bottom).

```{r ko-and-tax-ords, fig.height=pnas.2col*0.6}
set.seed(main.seed)
ord.list <- list() # creating empty list to store ordination objects
ord.plot.list <- list() # creating empty list to store PCoA plots

# do for both KOs and taxa, this loop will be used in most subsequent analyses
for (type in c("kos", "tax")) {
  # do for both forward and reverse reads
  for (readn in 1:2) {
    i <- paste(type, readn, sep = "-")
    # assign correct phyloseq object to `phy.obj`
    phy.obj <- eval(parse(text = paste("stool", readn, ".", type, ".phy", sep = "")))
    # ordinate with create method and distance metric
    ord <- ordinate(phy.obj, method = ord.method, distance = dist.method)
    # plot ordination and include data for correct number of axes
    ord.plot <- plot_ordination(phy.obj, ord, label = "smpl", axes = 1:nAxes) +
      ggtitle(
        paste(
          "read", readn, "s ",
          ifelse(type == "kos", "functional ", "taxonomy "),
          dist.method.name, " ", ord.method,
          sep = ""
        )
      ) +
      theme(plot.title = element_text(size = 13))
    # add ordination objects and plots to respective lists
    ord.list[[i]] <- ord
    ord.plot.list[[i]] <- ord.plot
    # save plot as an RDS file for easy loading in later
    saveRDS(
      ord.plot,
      file = file.path(
        saveDir,
        paste("ordPlot_read", readn, "s_",
              ifelse(type == "kos",
                     "functional_",
                     "taxonomy_"
              ),
              dist.method.name, "_", ord.method,
              ".rds",
              sep = ""
        )
      )
    )
  }
}
n <- length(ord.plot.list)
nCol <- floor(sqrt(n)) # this should make the panel layout as close to square as possible
# save ordination list as RDS for easy loading in later
saveRDS(ord.list, file = file.path(saveDir, "list_ords_pcoa.rds"))
# show ordinations
plot_grid(plotlist = ord.plot.list, ncol = 2)
```

## 2. KOs vs taxonomic Procrustes analyses

To infer whether there's a significant relationship between taxonomic and functional composition/diversity (as implied by that telling cluster of samples mentioned above), we ran a Procrustes analysis (`protest`{`vegan`}) comparing the functional and taxonomic ordinations for both forward and reverse reads sets. We also looked at the correlation between reads 1 and 2 within functional and taxonomic diversity. 

```{r ko-tax-procrustes, cache=TRUE, eval=TRUE}
set.seed(main.seed)
# Procrustes tests for all important comparisons
(r1v2.kvk <- protest(
  ord.list[["kos-1"]]$vectors, 
  ord.list[["kos-2"]]$vectors, 
  permutations = 9999
))
(r1v2.tvt <- protest(
  ord.list[["tax-1"]]$vectors, 
  ord.list[["tax-2"]]$vectors, 
  permutations = 9999
))
(r1v1.kvt <- protest(
  ord.list[["kos-1"]]$vectors, 
  ord.list[["tax-1"]]$vectors, 
  permutations = 9999
))
(r2v2.kvt <- protest(
  ord.list[["kos-2"]]$vectors, 
  ord.list[["tax-2"]]$vectors, 
  permutations = 9999
))

# Create a table of Procrustes tests results
prot.res <- data.frame(
  "read.comp" = c(rep("1 vs. 2", 2), "1 vs. 1", "2 vs. 2"),
  "type.comp" = c("KOs vs. KOs", "Taxa vs. Taxa", rep("KOs vs. Taxa", 2)),
  "R2" = round(c(r1v2.kvk$t0, r1v2.tvt$t0, r1v1.kvt$t0, r2v2.kvt$t0), 4),
  "sig" = c(r1v2.kvk$signif, r1v2.tvt$signif, r1v1.kvt$signif, r2v2.kvt$signif)
)

# write results table to html (makes it easy to convert to xlsx format with proper styling within Excel)
(prot.res.html <- prot.res %>%
    kable(align = c("c", "c", "r", "r")) %>%
    kable_styling(bootstrap_options = c("hover", "condensed"), full_width = F) %>%
    collapse_rows(columns = 1:2))
```

## 3. Variable selection

There are 58 meta-data covariates associated with these microbiome samples. In order to determine which of these variables are most important in influencing the diversity and composition of the microbiomes we ran a set of data reduction steps. We grouped the covariates into 5 categories: Child Behavior, Parenting, Environmental and Social Adveristy (ESA), Demography, and Gut-related. The last two categories contained covariates (such as age, ethnicity, and antibiotic use) that we know can influence the microbiome but that were not the focus of this particular study. Therefore, in subsequent analyes, we first accounted for the effects of the Demographic and Gut-related covariates that most strongly associated with the microbiome before the assessing the associations of the covariates within other categories,.

### Envfit

As a first passs to see which covariates within each category most strongly associated with the microbiome compostion, I applied the function `envfit`{`vegan`} to the same PCoA ordinations produced above (we ran two tests for each covariate category, using the 1st and 2nd PCoA axes, and then the third and fourth). Covariates that were deemed significant are colored in green, non-significant in red. 

```{r envfit-fig-gen, cache=FALSE, include=FALSE, eval=F}
#### Create Null objects ####
varSet.kos.12.plots <- NULL
varSet.tax.12.plots <- NULL
varSet.kos.34.plots <- NULL
varSet.tax.34.plots <- NULL
axes12.r2.data <- data.frame(NULL)
axes34.r2.data <- data.frame(NULL)
axis.pc.var <- data.frame()

#### Envfit analyses ####
for (varID in as.character(varIDs$varID)) { # for each covariate category
  grep.pattern <- paste("^", varID, sep = "")
  
  for (type in c("tax", "kos")) {
    phy.obj <- eval(parse(text = paste("stool1", type, "phy", sep = ".")))
    comm.tbl <- otu_table(phy.obj)
    metadata <- as(sample_data(phy.obj), "data.frame") %>% 
      as.data.table(keep.rownames = "SmplID")
    # remove samples with any NAs, otherwise `envfit` will fail
    keep.cols <- c("SmplID", grep(grep.pattern, names(metadata), value = T))
    varSet.data <- na.omit(metadata[, ..keep.cols])
    setkey(varSet.data, SmplID)
    if (type == "kos") {
      ord.i <- 1
    } else {
      ord.i <- 3
    }
    # read in ordination results from list saved above
    ord <- readRDS(file.path(saveDir, "list_ords_pcoa.rds"))[[ord.i]]
    # get ordination coordinates for each samples
    ord.points <- ord$vectors %>% as.data.table(keep.rownames = "SmplID")
    setkey(ord.points, SmplID)
    # combine covariate data with ordination coordinates
    # varSet.ordData <- cbind(varSet.data, varSet.ordPoints)
    varSet.ordData <- varSet.data[ord.points, nomatch = 0]
    ord.axes <- grep("Axis", names(varSet.ordData), value = T)
    varSet.ordPoints <- varSet.ordData[, ..ord.axes]
    # remove "_R1" from sample names
    varSet.ordData[, smpl := sub("_R1", "", SmplID)]
    # create the formula to be used in the `envfit` function
    # should look like `varSet.ordPoints ~ covar.1 + covar.2 + ... + covar.N`
    frm <- paste(
      "varSet.ordPoints ~",
      paste(sample(names(varSet.data)[-1]),
            collapse = " + "
      )
    )
    # run `envfit on ordinations of axes 1 & 2, and axes 3 & 4
    for (first.axis in c(1, 3)) {
      cat(
        paste(
          "varID='", varID, 
          "'; type='", type, 
          "'; first.axis=", first.axis, 
          sep = ""
        ),
        sep = "\n"
      )
      axes <- c(first.axis, first.axis + 1) # set first and second axes
      axes.names <- paste("Axis", axes, sep = ".") # set axes names (to match data.frame)
      # calculate variance explained for each axis
      var.exp <- round(ord$values$Eigenvalues / sum(ord$values$Eigenvalues) * 100, 1)
      # create a variance-explained data frame for use in plot
      var.exp.df <- data.frame(
        "type" = type,
        "first.axis" = first.axis,
        "axis1" = var.exp[first.axis],
        "axis2" = var.exp[first.axis + 1]
      )
      # add this variance-explained df to a larger one for all plots
      axis.pc.var <- rbind(axis.pc.var, var.exp.df)
      # run `envfit` analysis
      set.seed(main.seed)
      varSet.envfit <- envfit(
        as.formula(frm),
        varSet.data,
        perm = 9999,
        choices = axes
      )
      # extract results for vectors
      varSet.vectors <- data.frame(
        as.data.frame(varSet.envfit$vectors$arrows),
        "r2" = varSet.envfit$vectors$r,
        "p" = varSet.envfit$vectors$pvals,
        "name" = mgsub(
          vars.legend$var.name,
          vars.legend$description,
          row.names(varSet.envfit$vectors$arrows)
        ),
        stringsAsFactors = F
      )
      # extract results for factors
      varSet.factors <- data.frame(
        "r2" = varSet.envfit$factors$r,
        "p" = varSet.envfit$factors$pvals,
        "name" = mgsub(
          vars.legend$var.name,
          vars.legend$description,
          names(varSet.envfit$factors$r)
        )
      )
      # combine vector and factor results and clean up
      varSet.vecs.facs <- rbind(varSet.vectors[, -c(1, 2)], varSet.factors)
      varSet.vecs.facs$var <- row.names(varSet.vecs.facs)
      varSet.vecs.facs$t <- type
      row.names(varSet.vecs.facs) <- NULL
      # append combined results to the proper data.frame for choosing covariates later
      if (first.axis == 1) {
        axes12.r2.data <- rbind(axes12.r2.data, varSet.vecs.facs)
      } else {
        axes34.r2.data <- rbind(axes34.r2.data, varSet.vecs.facs)
      }
      # create the base plot with sample points on ordination
      envfit.plot <- ggplot(
        data = varSet.ordData,
        aes_string(x = axes.names[1], y = axes.names[2])
      ) +
        geom_point(shape = 1, size = inner.pt.sz) +
        scale_color_manual(values = unname(c(cb.pal[c("red", "green")]))) +
        labs(
          title = mgsub(varIDs$varID, varIDs$name, varID) # ,
          # x=paste(axes.names[1], " [", var.exp[axes[1]], "%]", sep=""),
          # y=paste(axes.names[2], " [", var.exp[axes[2]], "%]", sep="")
        ) +
        theme(
          legend.position = "none",
          plot.title = element_text(size = 10),
          axis.title = element_blank()
        )
      # add vectors, scaled appropriately (taken from experimenting with default 
      # `plot(envfit)`), if applicable
      if (nrow(varSet.vectors) > 0) {
        if (ordiArrowMul(varSet.envfit) < 1) {
          varSet.vectors$scale1 <- varSet.vectors[, axes.names[1]] *
            sqrt(varSet.vectors$r2) *
            ordiArrowMul(varSet.envfit)
          varSet.vectors$scale2 <- varSet.vectors[, axes.names[2]] *
            sqrt(varSet.vectors$r2) *
            ordiArrowMul(varSet.envfit)
        } else {
          varSet.vectors$scale1 <- varSet.vectors[, axes.names[1]] *
            sqrt(varSet.vectors$r2) /
            ordiArrowMul(varSet.envfit)
          varSet.vectors$scale2 <- varSet.vectors[, axes.names[2]] *
            sqrt(varSet.vectors$r2) /
            ordiArrowMul(varSet.envfit)
        }
        envfit.plot <- envfit.plot +
          geom_segment(
            data = varSet.vectors,
            x = 0,
            y = 0,
            aes(
              xend = scale1,
              yend = scale2
            ), # /12
            arrow = arrow(length = unit(0.02, "npc")),
            size = 0.5,
            color = "blue"
          ) +
          geom_text_repel(
            data = varSet.vectors,
            aes(
              x = scale1,
              y = scale2, # /12
              label = name,
              color = ifelse(p < 0.05, "sig.", "non-sig.")
            ),
            size = lab.size
          )
      }
      # add factor centroids, if applicable
      if (nrow(varSet.factors) > 0) {
        centroids <- as.data.frame(varSet.envfit$factors$centroids)
        centroids$var <- substr(row.names(centroids), 1, nchar(row.names(centroids)) - 1)
        centroids$centroid.lab <- mgsub(
          vars.legend$var.name,
          vars.legend$description,
          row.names(centroids)
        )
        centroid.dat <- merge(
          centroids, 
          varSet.factors, 
          by.x = "var", 
          by.y = 0, 
          all.x = TRUE
        )
        centroid.dat$sig <- ifelse(centroid.dat$p < 0.05, "sig.", "non-sig.")
        envfit.plot <- envfit.plot +
          geom_point(
            data = centroid.dat,
            aes_string(
              x = axes.names[1],
              y = axes.names[2]
            ),
            shape = 3,
            size = inner.pt.sz,
            color = "blue"
          ) +
          geom_text_repel(
            data = centroid.dat,
            aes_string(
              x = axes.names[1],
              y = axes.names[2],
              label = "centroid.lab",
              color = "sig"
            ),
            size = lab.size
          )
      }
      # add plot to appropriate list
      if (type == "kos") {
        if (first.axis == 1) {
          varSet.kos.12.plots[[varID]] <- envfit.plot
        } else {
          varSet.kos.34.plots[[varID]] <- envfit.plot
        }
      } else {
        if (first.axis == 1) {
          varSet.tax.12.plots[[varID]] <- envfit.plot
        } else {
          varSet.tax.34.plots[[varID]] <- envfit.plot
        }
      }
    }
  }
}
axis.pc.var <- unique(axis.pc.var) # deduplicate

#### Save Everything ####
saveRDS(varSet.kos.12.plots, file = file.path(saveDir, "list_envfitPlots_varSets_kos_axes12.rds"))
saveRDS(varSet.tax.12.plots, file = file.path(saveDir, "list_envfitPlots_varSets_tax_axes12.rds"))
saveRDS(varSet.kos.34.plots, file = file.path(saveDir, "list_envfitPlots_varSets_kos_axes34.rds"))
saveRDS(varSet.tax.34.plots, file = file.path(saveDir, "list_envfitPlots_varSets_tax_axes34.rds"))
saveRDS(axis.pc.var, file = file.path(saveDir, "df_envfit_axis_pc_var.rds"))

saveRDS(axes12.r2.data, file = file.path(saveDir, "df_envfit_axes12_r2_data.rds"))
saveRDS(axes34.r2.data, file = file.path(saveDir, "df_envfit_axes34_r2_data.rds"))

#### Check it out ####
axes12.r2.data <- readRDS(file.path(saveDir, "df_envfit_axes12_r2_data.rds"))
axes34.r2.data <- readRDS(file.path(saveDir, "df_envfit_axes34_r2_data.rds"))
r2.data <- rbind(axes12.r2.data, axes34.r2.data)
sig.r2.data <- subset(r2.data, p < 0.05) %>% droplevels()
sig.r2.data <- sig.r2.data[order(sig.r2.data$r2, decreasing = T), ]
kos.sig.vars <- lapply(varIDs$varID, function(var.cat) {
  subset(sig.r2.data, t == "kos" & grepl(var.cat, var))$var %>% unique()
})
names(kos.sig.vars) <- varIDs$varID
tax.sig.vars <- lapply(varIDs$varID, function(var.cat) {
  subset(sig.r2.data, t == "tax" & grepl(var.cat, var))$var %>% unique()
})
names(tax.sig.vars) <- varIDs$varID
analysis.terms <- list("kos" = kos.sig.vars, "tax" = tax.sig.vars)
# this is the list of covariate that will be used in all subsequent analyses
saveRDS(analysis.terms, file = file.path(saveDir, "list_analysis_terms_envfit.rds"))
```

```{r envfit-all-sig-vars}
all.terms <- readRDS(file.path(saveDir, "list_analysis_terms_envfit.rds"))
cat("Significant terms based on functional composition", sep = "\n")
(kos.terms <- lapply(
  all.terms$kos, 
  function(x) mgsub(vars.legend$var.name, vars.legend$description, x)
  ))
cat("Significant terms based on taxonomic composition", sep = "\n")
(tax.terms <- lapply(
  all.terms$tax, 
  function(x) mgsub(vars.legend$var.name, vars.legend$description, x)
  ))

```

#### Taxonomic Diversity

```{r envfit-tax, fig.height=pnas.max.ht, fig.height=20}
# show `envfit` results for taxonomic-based diversity
varSet.tax.12.plots <- readRDS(
  file = file.path(saveDir, "list_envfitPlots_varSets_tax_axes12.rds")
)
varSet.tax.34.plots <- readRDS(
  file = file.path(saveDir, "list_envfitPlots_varSets_tax_axes34.rds")
)
axis.pc.var <- readRDS(file.path(saveDir, "df_envfit_axis_pc_var.rds")) %>% as.data.table()
pc1 <- axis.pc.var[type == "tax" & first.axis == 1, axis1] %>% sprintf("%.1f", .)
pc2 <- axis.pc.var[type == "tax" & first.axis == 1, axis2] %>% sprintf("%.1f", .)
varSet.tax.12.grob <- arrangeGrob(
  grobs = varSet.tax.12.plots,
  ncol = 1,
  top = "PCoA Axes 1 and 2",
  bottom = paste("Axis.1 [", pc1, "%]", sep = ""),
  left = paste("Axis.2 [", pc2, "%]", sep = "")
)
pc3 <- axis.pc.var[type == "tax" & first.axis == 3, axis1] %>% sprintf("%.1f", .)
pc4 <- axis.pc.var[type == "tax" & first.axis == 3, axis2] %>% sprintf("%.1f", .)
varSet.tax.34.grob <- arrangeGrob(
  grobs = varSet.tax.34.plots,
  ncol = 1,
  top = "PCoA Axes 3 and 4",
  bottom = paste("Axis.3 [", pc3, "%]", sep = ""),
  left = paste("Axis.4 [", pc4, "%]", sep = "")
)
varSet.tax.grob <- arrangeGrob(varSet.tax.12.grob, varSet.tax.34.grob, nrow = 1)
grid.arrange(varSet.tax.grob)
```

#### Functional Diversity

```{r envfit-kos, fig.height=pnas.max.ht, fig.height=20}
# show `envfit` results for functional-based diversity
varSet.kos.12.plots <- readRDS(
  file = file.path(saveDir, "list_envfitPlots_varSets_kos_axes12.rds")
)
varSet.kos.34.plots <- readRDS(
  file = file.path(saveDir, "list_envfitPlots_varSets_kos_axes34.rds")
)
axis.pc.var <- readRDS(file.path(saveDir, "df_envfit_axis_pc_var.rds")) %>% as.data.table()
pc1 <- axis.pc.var[type == "kos" & first.axis == 1, axis1] %>% sprintf("%.1f", .)
pc2 <- axis.pc.var[type == "kos" & first.axis == 1, axis2] %>% sprintf("%.1f", .)
varSet.kos.12.grob <- arrangeGrob(
  grobs = varSet.kos.12.plots,
  ncol = 1,
  top = "PCoA Axes 1 and 2",
  bottom = paste("Axis.1 [", pc1, "%]", sep = ""),
  left = paste("Axis.2 [", pc2, "%]", sep = "")
)
pc3 <- axis.pc.var[type == "kos" & first.axis == 3, axis1] %>% sprintf("%.1f", .)
pc4 <- axis.pc.var[type == "kos" & first.axis == 3, axis2] %>% sprintf("%.1f", .)
varSet.kos.34.grob <- arrangeGrob(
  grobs = varSet.kos.34.plots,
  ncol = 1,
  top = "PCoA Axes 3 and 4",
  bottom = paste("Axis.3 [", pc3, "%]", sep = ""),
  left = paste("Axis.4 [", pc4, "%]", sep = "")
)
varSet.kos.grob <- arrangeGrob(varSet.kos.12.grob, varSet.kos.34.grob, nrow = 1)
grid.arrange(varSet.kos.grob)
```

## 4. Constrained coorespondence analysis

For the following analyses, we used a constrained coorespondence analysis (CCA) followed by PERMANOVA to test our hypotheses regarding the relationship between the selected covariates and the microbiome. We used CCA because it allows us to account for the variance explained by Demographic and Gut-related covariates (which we know from previous research may have an important effect, but we don't want to focus on here), and then test the relationships of our selected covariates and the remaining variance.

###  ESA by Parenting


Here, we generated an original model with the form: `distance matrix ~ (esa_var1 + esa_var2 + ...) * (parent_var1 + parent_var2 + ...)` and then ran `ordistep` on this model to select for the best main effects and interaction terms.

```{r cca-esaByPar-analysis, cache=FALSE, eval=TRUE}
set.seed(main.seed)
redo.ordistep <- FALSE
# name of file containing list of `envfit`-selected covariates
terms.file <- file.path(saveDir, "list_analysis_terms_envfit.rds")
type <- "kos"
for (type in c("tax", "kos")) {
  # name file that will be used for writing the `ordistep` results
  ordistep.file <- file.path(saveDir, paste("ordistep_ESAbyPar_", type, ".rds", sep = ""))
  # load all selected covariates
  analysis.terms <- readRDS(terms.file)[[type]]
  # a list of the selected ESA covariate(s)
  focal.vars <- analysis.terms[["esa_"]]
  # a list of the selected parenting covariate(s)
  mod.vars <- analysis.terms[["parent_"]]
  # a list of the selected demography and gut-history covariate(s)
  cond.vars <- c(
    analysis.terms[["demo_"]], 
    analysis.terms[["gut_"]], 
    analysis.terms[["Diet_"]]
  )
  cond.vars <- cond.vars[!is.na(cond.vars)] # make sure no NAs
  
  # assign correct phyloseq object to `phy.obj`
  phy.obj <- eval(parse(text = paste("stool1", type, "phy", sep = ".")))
  # extract KO/taxa community table and metadata table from `phy.obj`
  comm.tbl <- otu_table(phy.obj)
  metadata <- as(sample_data(phy.obj), "data.frame")
  # remove samples with NAs in covars (or CCA will fail)
  var.data <- na.omit(metadata[, c(focal.vars, mod.vars, cond.vars), drop = F])
  # also drop those samples from community table
  var.comm <- comm.tbl[row.names(var.data), ]
  
  # if there is only one focal covar and one moderating var, run CCA analysis with conditions
  # and no model selection
  if (length(focal.vars) == 1 & length(mod.vars) == 1) {
    frm <- paste("var.comm ~", focal.vars, "*", mod.vars)
    if (length(cond.vars) > 0) {
      frm <- paste(frm, " + Condition(",
                   paste(cond.vars, collapse = " + "), ")",
                   sep = ""
      )
    }
    cca.obj <- cca(as.formula(frm), data = var.data)
  } else {
    # if there are multiple focal/moderatings covars, run model selection (without conditions)
    # using `ordistep`
    if (redo.ordistep) {
      # create appropriate formula for a running `cca`:
      # `var.comm ~ focal.var.1 + focal.var2 + ... + focal.var.N +
      #             mod.var.1 + mod.var.2 + ... + mod.var.N`
      frm0 <- paste("var.comm ~ (",
                    paste(focal.vars, collapse = " + "),
                    ") * (",
                    paste(mod.vars, collapse = " + "),
                    ")",
                    sep = ""
      )
      
      # create full CCA model
      cca.obj0 <- cca(as.formula(frm0), data = var.data)
      # run model selection to possibly reduce variables
      # set.seed(main.seed)
      cca.ordi <- ordistep(cca.obj0, direction = "backward", trace = TRUE)
      saveRDS(cca.ordi, file = ordistep.file)
    } else {
      cca.ordi <- readRDS(ordistep.file)
    }
    # create new formula with covars selected by `ordistep` and
    # now include condition covars (demography and gut-history) in formula
    frm <- paste(
      "var.comm ~",
      as.character(cca.ordi$call$formula)[3]
    )
    if (length(cond.vars) > 0) {
      frm <- paste(frm, " + Condition(",
                   paste(cond.vars, collapse = " + "), ")",
                   sep = ""
      )
    }
    # run CCA analysis (this one will be used for assessing significance)
    cca.obj <- cca(as.formula(frm), data = var.data)
  }
  saveRDS(
    cca.obj, 
    file = file.path(saveDir, paste("cca_ESAbyPar_", type, ".rds", sep = ""))
  )
  # run PERMANOVA on CCA object to assess signficance
  cca.aov.term <- anova(cca.obj, by = "terms", permutations = 9999, parallel = nproc)
  print(cca.aov.term) # print PERMANOVA results and save them below
  saveRDS(
    cca.aov.term,
    file = file.path(saveDir, paste("anova_cca", type, "ESAByPar_term.rds", sep = "_"))
  )
}
```

```{r cca-esaByPar-aovs, results='asis'}
# create nice tables for looking as PERMANOVA results
tax.aov.full <- readRDS(file.path(saveDir, "anova_cca_tax_ESAbyPar_term.rds"))
tax.aov <-  as.data.frame(tax.aov.full)
kos.aov.full <- readRDS(file.path(saveDir, "anova_cca_kos_ESAbyPar_term.rds"))
kos.aov <- as.data.frame(kos.aov.full)
row.names(tax.aov) <- mgsub(vars.legend$var.name, vars.legend$description, row.names(tax.aov))
tax.aov[, "sig"] <- ifelse(tax.aov$`Pr(>F)` < 0.05 & !is.na(tax.aov$`Pr(>F)`), "*", "")
tax.aov <- tibble::rownames_to_column(tax.aov, var = "covar")
cat(attributes(tax.aov.full)$heading[2], sep = "\n")
(tax.html <- tax.aov %>%
    kable(align = c(rep("r", ncol(tax.aov) - 1), "l"), digits = 3, caption = "Taxonomic diversity") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
    column_spec(1, border_right = T))

row.names(kos.aov) <- mgsub(vars.legend$var.name, vars.legend$description, row.names(kos.aov))
kos.aov[, "sig"] <- ifelse(kos.aov$`Pr(>F)` < 0.05 & !is.na(kos.aov$`Pr(>F)`), "*", "")
cat(attributes(kos.aov.full)$heading[2], sep = "\n")
(kos.html <- kos.aov %>%
    tibble::rownames_to_column(var = "covar") %>%
    kable(align = c(rep("r", 5), "l"), digits = 3, caption = "Functional diversity") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
    column_spec(1, border_right = T))
```

```{r cca-esaByPar-cap, eval=TRUE}
esaByPar.cap <- "Constrained coorespondance analysis (CCA) ordinations of taxonomic (A & B) and functional diversity (C & D). Each point represents a single fecal sample. (A) Significant association between taxonomic diversity and the ESA covariate, LEC Turmoil. Points are colored by individual LEC Turmoil scores with purple indicating low and green indicating high. (B) Significant interaction between Income to Needs, Parent-Child Dysfunction and taxonomic diversity of the microbiome. The color of the outer ring of each point indicates the individual Income to Needs score of the child, where blue indicates low and red indicates high, and the inner circle of each point is colored according to the parent-reported Parent-Child Dysfunction score, where white indicates low dysfunction and black indicate high dysfunction. (C) Significant interaction between LEC Turmoil, Parent-Child Dysfunction, and functional diversity of the microbiome. The outer ring of each point is colored by the individual LEC Turmoil scores as in panel A, and the inner circle is colored by the Parent-Child Dysfunction score as in panel B. (D) Significant interaction between Income to Needs, Parent-Child Dysfunction, and functional diversity of the microbiome. Points are colored in an identical manner to panel B."
```

```{r cca-esaByPar-fig-gen, fig.height=pnas.max.ht/2}
# create ordination figures for ESA CCA results showing only significant covars

# assign correct phyloseq objects to new variable names
tax.obj <- stool1.tax.phy
kos.obj <- stool1.kos.phy
# read in saved CCA results
(tax.cca <- readRDS(file.path(saveDir, "cca_ESAbyPar_tax.rds")))
(kos.cca <- readRDS(file.path(saveDir, "cca_ESAbyPar_kos.rds")))
# read in saved PERMANOVA results
tax.aov <- readRDS(file.path(saveDir, "anova_cca_tax_ESAbyPar_term.rds"))
kos.aov <- readRDS(file.path(saveDir, "anova_cca_kos_ESAbyPar_term.rds"))

# extract significant ESA covars from taxonomic diversity PERMANOVA table
tax.aov <- tax.aov[complete.cases(tax.aov), ]
tax.sig.labs <- row.names(tax.aov[tax.aov$`Pr(>F)` <= 0.05, ])
if (any(grepl(":", tax.sig.labs))) {
  split.labs <- strsplit(tax.sig.labs[grepl(":", tax.sig.labs)], ":") %>% unlist()
  tax.sig.vars <- unique(c(split.labs, tax.sig.labs))
} else {
  tax.sig.vars <- tax.sig.labs
}
# extract significant ESA covars from functional diversity PERMANOVA table
kos.aov <- kos.aov[complete.cases(kos.aov), ]
kos.sig.labs <- row.names(kos.aov[kos.aov$`Pr(>F)` <= 0.05, ])
if (any(grepl(":", kos.sig.labs))) {
  split.labs <- strsplit(kos.sig.labs[grepl(":", kos.sig.labs)], ":") %>% unlist()
  kos.sig.vars <- unique(c(split.labs, kos.sig.labs))
} else {
  kos.sig.vars <- kos.sig.labs
}

# create plots
metadata <- as(sample_data(kos.obj), "data.frame")
plot.axes <- c(1, 2)
tax.sites <- as.data.frame(tax.cca$CCA$u[, plot.axes]) # sample points
tax.axes <- names(tax.sites)[plot.axes]
tax.sites$smpl <- sub("_R1", "", row.names(tax.sites)) # cleanup sample names
tax.sites <- merge(tax.sites, metadata, by = "smpl", all.x = T)
# extract vector data for significant covariates
tax.spec <- as.data.frame(tax.cca$CCA$v)
tax.spec[, "kos"] <- row.names(tax.spec)
tax.scale <- ordiArrowMul(tax.cca$CCA$u[, plot.axes])
tax.arws <- as.data.frame(tax.cca$CCA$biplot[, plot.axes] / tax.scale)
tax.arws <- tax.arws[tax.sig.vars, ]
tax.arws$var <- mgsub(
  vars.legend$var.name,
  vars.legend$description,
  row.names(tax.arws)
)

# LEC Turmoil plot
lec.tur.tax0 <- ggplot(data = tax.sites, aes_string(tax.axes[1], tax.axes[2])) +
  geom_segment(
    data = tax.arws, x = 0, y = 0,
    aes_string(
      xend = tax.axes[1],
      yend = tax.axes[2]
    ),
    arrow = arrow(length = unit(0.02, "npc")),
    color = "royalblue4", size = seg.size
  ) +
  geom_point(
    aes_string(color = row.names(tax.arws)[1]),
    shape = 19, size = outer.pt.sz
  ) +
  geom_point(
    aes_string(fill = row.names(tax.arws)[2]),
    shape = 21, size = inner.pt.sz, color = "white"
  ) +
  geom_text_repel(
    data = tax.arws,
    aes_string(
      x = tax.axes[1],
      y = tax.axes[2],
      label = "var"
    ),
    size = lab.size,
    color = "royalblue4"
  ) +
  scale_color_gradientn(
    name = vars.legend[var.name == row.names(tax.arws)[1]]$description, 
    colors = blue_to_red
  ) +
  scale_fill_gradient(
    name = gsub(" ", "\n", vars.legend[var.name == row.names(tax.arws)[2]]$description),
    low = "white", 
    high = "black"
  ) +
  labs(title = "A", subtitle = "Taxonomic diversity") +
  theme(legend.position = "bottom")
leg1 <- get_legend(lec.tur.tax0)
lec.tur.tax <- lec.tur.tax0 + theme(legend.position = 'none')

kos.sites <- as.data.frame(kos.cca$CCA$u[, plot.axes]) # sample points
kos.axes <- names(kos.sites)[plot.axes]
kos.sites$smpl <- sub("_R1", "", row.names(kos.sites)) # cleanup sample names
kos.sites <- merge(kos.sites, metadata, by = "smpl", all.x = T)
# extract vector data for significant covariates
kos.spec <- as.data.frame(kos.cca$CCA$v)
kos.spec[, "kos"] <- row.names(kos.spec)
kos.scale <- ordiArrowMul(kos.cca$CCA$u[, plot.axes])
kos.arws <- as.data.frame(kos.cca$CCA$biplot[, plot.axes] / kos.scale)
kos.arws <- kos.arws[kos.sig.vars, ]
kos.arws$var <- mgsub(
  vars.legend$var.name,
  vars.legend$description,
  row.names(kos.arws)
)
# LEC Turmoil by Parent-Child Dysfunction plot
tur.pcd.kos0 <- ggplot(data = kos.sites, aes_string(kos.axes[1], kos.axes[2])) +
  geom_segment(
    data = kos.arws, 
    x = 0, 
    y = 0,
    aes_string(
      xend = kos.axes[1],
      yend = kos.axes[2]
    ),
    arrow = arrow(length = unit(0.02, "npc")),
    color = "royalblue4", size = seg.size
  ) +
  geom_point(
    aes_string(color = row.names(kos.arws)[1]),
    shape = 19, size = outer.pt.sz
  ) +
  geom_point(
    aes_string(fill = row.names(kos.arws)[2]),
    shape = 21, size = inner.pt.sz, color = "white"
  ) +
  geom_text_repel(
    data = kos.arws,
    aes_string(
      x = kos.axes[1],
      y = kos.axes[2],
      label = "var"
    ),
    size = lab.size,
    color = "royalblue4"
  ) +
  scale_color_gradientn(name = kos.arws$var[1], colors = purple_to_green) +
  scale_fill_gradient(
    name = sub("Child ", "Child\n", kos.arws$var[2]), 
    low = "white", 
    high = "black"
  ) +
  labs(title = "B", subtitle = "Functional diversity") +
  guides(fill = "none") +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom"
  )
leg2 <- get_legend(tur.pcd.kos0)
tur.pcd.kos <- tur.pcd.kos0 + theme(legend.position = "none")

# create panelled figure
plots.grob <- plot_grid(lec.tur.tax, tur.pcd.kos, nrow = 1)
wds <- c(0.20, 0.25, 0.18, 0.20, 0.17)
sum(wds)
legs.grob <- plot_grid(
  NULL, leg1, NULL, leg2, NULL, 
  rel_widths = wds,
  nrow = 1
)
figure <- plot_grid(plots.grob, legs.grob, ncol = 1, rel_heights = c(0.9, 0.1))
print(figure)
```

### Child Behavior by Parenting

```{r cca-chbByParent-analysis, cache=FALSE, eval=TRUE}
set.seed(main.seed)
redo.ordistep <- FALSE
terms.file <- file.path(saveDir, "list_analysis_terms_envfit.rds")
for (type in c("tax")) {
  ordistep.file <- file.path(saveDir, paste("ordistep_chbByPar_", type, ".rds", sep = ""))
  analysis.terms <- readRDS(terms.file)[[type]]
  focal.vars <- analysis.terms[["child_behav_"]]
  mod.vars <- analysis.terms[["parent_"]]
  cond.vars <- c(
    analysis.terms[["demo_"]], 
    analysis.terms[["gut_"]], 
    analysis.terms[["Diet_"]]
  )
  cond.vars <- cond.vars[!is.na(cond.vars)]
  phy.obj <- eval(parse(text = paste("stool1", type, "phy", sep = ".")))
  comm.tbl <- otu_table(phy.obj)
  metadata <- as(sample_data(phy.obj), "data.frame")
  var.data <- na.omit(metadata[, c(focal.vars, mod.vars, cond.vars), drop = F])
  var.comm <- comm.tbl[row.names(var.data), ]
  if (length(focal.vars) == 1 & length(mod.vars) == 1) {
    frm <- paste("var.comm ~", focal.vars, "*", mod.vars)
    if (length(cond.vars) > 0) {
      frm <- paste(frm, " + Condition(",
                   paste(cond.vars, collapse = " + "), ")",
                   sep = ""
      )
    }
    cca.obj <- cca(as.formula(frm), data = var.data)
  } else {
    if (redo.ordistep) {
      frm0 <- paste("var.comm ~ (",
                    paste(focal.vars, collapse = " + "),
                    ") * (",
                    paste(mod.vars, collapse = " + "),
                    ")",
                    sep = ""
      )
      print(frm0)
      cca.obj0 <- cca(as.formula(frm0), data = var.data)
      cca.ordi <- ordistep(cca.obj0, direction = "backward", trace = TRUE)
      saveRDS(cca.ordi, file = ordistep.file)
    } else {
      cca.ordi <- readRDS(ordistep.file)
    }
    frm <- paste(
      "var.comm ~",
      as.character(cca.ordi$call$formula)[3]
    )
    if (length(cond.vars) > 0) {
      frm <- paste(frm, " + Condition(",
                   paste(cond.vars, collapse = " + "), ")",
                   sep = ""
      )
    }
    cca.obj <- cca(as.formula(frm), data = var.data)
  }
  saveRDS(
    cca.obj, 
    file = file.path(saveDir, paste("cca_chbByPar_", type, ".rds", sep = ""))
  )
  # cca.aov.marg <- anova(cca.obj, by="margin", permutations=999, parallel=nproc)
  cca.aov.term <- anova(cca.obj, by = "terms", permutations = 9999, parallel = nproc)
  # saveRDS(cca.aov.marg, file=file.path(saveDir, paste("anova_cca", type, "chbByPar_marg.rds", sep="_")))
  saveRDS(
    cca.aov.term,
    file = file.path(saveDir, paste("anova_cca", type, "chbByPar_term.rds", sep = "_"))
  )
}
```

```{r cca-chbByPar-aovs, results='asis'}
tax.aov.full <- readRDS(file.path(saveDir, "anova_cca_tax_chbByPar_term.rds"))
tax.aov <-  as.data.frame(tax.aov.full)
kos.aov.full <- readRDS(file.path(saveDir, "anova_cca_kos_chbByPar_term.rds"))
kos.aov <- as.data.frame(kos.aov.full)
row.names(tax.aov) <- mgsub(vars.legend$var.name, vars.legend$description, row.names(tax.aov))
tax.aov[, "sig"] <- ifelse(tax.aov$`Pr(>F)` < 0.05 & !is.na(tax.aov$`Pr(>F)`), "*", "")
tax.aov <- rownames_to_column(tax.aov, var = "covar")
cat(attributes(tax.aov.full)$heading[2], sep = "\n")
(tax.html <- tax.aov %>%
    kable(align = c(rep("r", ncol(tax.aov) - 1), "l"), digits = 3, caption = "Taxonomic diversity") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
    column_spec(1, border_right = T))

row.names(kos.aov) <- mgsub(vars.legend$var.name, vars.legend$description, row.names(kos.aov))
kos.aov[, "sig"] <- ifelse(kos.aov$`Pr(>F)` < 0.05 & !is.na(kos.aov$`Pr(>F)`), "*", "")
cat(attributes(kos.aov.full)$heading[2], sep = "\n")
(kos.html <- kos.aov %>%
    rownames_to_column(var = "covar") %>%
    kable(align = c(rep("r", 5), "l"), digits = 3, caption = "Functional diversity") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
    column_spec(1, border_right = T))
```

```{r cca-chbByPar-cap, eval=FALSE}
chbByPar.cap <- "Constrained coorespondance analysis (CCA) ordinations of taxonomic (A) and functional diversity (B & C). Each point represents a single fecal sample. (A) Significant association between taxonomic diversity and the Child Behavior covariate, Depressive Problems. Points are colored by individual Depressive Problems scores with blue indicating low and red indicating high. (B) Significant interaction between Inhibitory Control, Parent-Child Dysfunction and functional diversity of the microbiome. The color of the outer ring of each point indicates the individual Inhibitory Control score of the child, where purple indicates low and red indicates high, and the inner circle of each point is colored according to the parent-reported Parent-Child Dysfunction score, where white indicates low dysfunction and black indicate high dysfunction. (C) Significant interaction between Depressive Problems, Parent-Child Dysfunction, and functional diversity of the microbiome. The outer ring of each point is colored by the individual Depressive Problems scores as in panel A, and the inner circle is colored by the Parent-Child Dysfunction score as in panel B."
```

```{r cca-chbByPar-fig-gen, fig.height=pnas.max.ht/1.5, fig.cap=chbByPar.cap, eval=FALSE}
tax.obj <- stool1.tax.phy
kos.obj <- stool1.kos.phy
(tax.cca <- readRDS(file.path(saveDir, "cca_chbByPar_tax.rds")))
(kos.cca <- readRDS(file.path(saveDir, "cca_chbByPar_kos.rds")))
tax.aov <- readRDS(file.path(saveDir, "anova_cca_tax_chbByPar_term.rds"))
kos.aov <- readRDS(file.path(saveDir, "anova_cca_kos_chbByPar_term.rds"))

tax.aov <- tax.aov[complete.cases(tax.aov), ]
tax.sig.labs <- row.names(tax.aov[tax.aov$`Pr(>F)` <= 0.05, ])
if (any(grepl(":", tax.sig.labs))) {
  split.labs <- strsplit(tax.sig.labs[grepl(":", tax.sig.labs)], ":") %>% unlist()
  tax.sig.vars <- unique(c(split.labs, tax.sig.labs))
} else {
  tax.sig.vars <- tax.sig.labs
}
kos.aov <- kos.aov[complete.cases(kos.aov), ]
kos.sig.labs <- row.names(kos.aov[kos.aov$`Pr(>F)` <= 0.05, ])
if (any(grepl(":", kos.sig.labs))) {
  split.labs <- strsplit(kos.sig.labs[grepl(":", kos.sig.labs)], ":") %>% unlist()
  kos.sig.vars <- unique(c(split.labs, kos.sig.labs))
} else {
  kos.sig.vars <- kos.sig.labs
}

metadata <- as(sample_data(kos.obj), "data.frame")
plot.axes <- c(1, 2)
kos.sites <- as.data.frame(kos.cca$CCA$u[, plot.axes])
kos.axes <- names(kos.sites)[plot.axes]
kos.sites$smpl <- sub("_R1", "", row.names(kos.sites))
kos.sites <- merge(kos.sites, metadata, by = "smpl", all.x = T)
kos.spec <- as.data.frame(kos.cca$CCA$v)
kos.spec[, "kos"] <- row.names(kos.spec)
kos.scale <- ordiArrowMul(kos.cca$CCA$u[, plot.axes])
kos.arws <- as.data.frame(kos.cca$CCA$biplot[, plot.axes] / kos.scale)
kos.arws <- kos.arws[kos.sig.vars, ]
kos.arws$var <- mgsub(
  vars.legend$var.name,
  vars.legend$description,
  row.names(kos.arws)
)

imp.kos.arws <- kos.arws[4, ]
imp.kos0 <- ggplot(data = kos.sites, aes_string(kos.axes[1], kos.axes[2])) +
  geom_segment(
    data = imp.kos.arws, x = 0, y = 0,
    aes_string(
      xend = kos.axes[1],
      yend = kos.axes[2]
    ),
    arrow = arrow(length = unit(0.02, "npc")),
    color = "royalblue4", size = seg.size
  ) +
  geom_point(
    aes_string(color = row.names(imp.kos.arws)[1]),
    shape = 19, size = outer.pt.sz
  ) +
  geom_text_repel(
    data = imp.kos.arws,
    aes_string(
      x = kos.axes[1],
      y = kos.axes[2],
      label = "var"
    ),
    size = lab.size,
    color = "royalblue4"
  ) +
  scale_color_viridis_c(name = gsub(" ", "\n", imp.kos.arws$var)) +
  labs(title = "A", subtitle = "Functional diversity") +
  theme(legend.position = "bottom", legend.box.just = 1)
leg1 <- get_legend(imp.kos0)
imp.kos <- imp.kos0 + theme(legend.position = "none")

ic.pcd.kos.arws <- kos.arws[c(1, 2, 5), ]
ic.pcd.kos0 <- ggplot(data = kos.sites, aes_string(kos.axes[1], kos.axes[2])) +
  geom_segment(
    data = ic.pcd.kos.arws, x = 0, y = 0,
    aes_string(
      xend = kos.axes[1],
      yend = kos.axes[2]
    ),
    arrow = arrow(length = unit(0.02, "npc")),
    color = "royalblue4", size = seg.size
  ) +
  geom_point(
    aes_string(color = row.names(ic.pcd.kos.arws)[1]),
    shape = 19, size = outer.pt.sz
  ) +
  geom_point(
    aes_string(fill = row.names(ic.pcd.kos.arws)[2]),
    shape = 21, size = inner.pt.sz, color = "white"
  ) +
  geom_text_repel(
    data = ic.pcd.kos.arws,
    aes_string(
      x = kos.axes[1],
      y = kos.axes[2],
      label = "var"
    ),
    size = lab.size,
    color = "royalblue4"
  ) +
  scale_color_gradientn(
    name = gsub(" ", "\n", ic.pcd.kos.arws$var[1]), 
    colors = blue_to_red
    ) +
  scale_fill_gradient(
    name = gsub(" ", "\n", ic.pcd.kos.arws$var[2]), 
    low = "white", 
    high = "black"
    ) +
  labs(title = "B", subtitle = "Functional diversity") +
  theme(legend.position = "bottom", axis.title.y = element_blank(), legend.box.just = 1)
leg2.1 <- get_legend(ic.pcd.kos0)
leg2.2 <- get_legend(ic.pcd.kos0 + theme(legend.box = "vertical", legend.box.just = 1))
ic.pcd.kos <- ic.pcd.kos0 + theme(legend.position = "none")

dep.pcd.kos.arws <- kos.arws[c(3, 2, 6), ]
dep.pcd.kos0 <- ggplot(data = kos.sites, aes_string(kos.axes[1], kos.axes[2])) +
  geom_segment(
    data = dep.pcd.kos.arws, x = 0, y = 0,
    aes_string(
      xend = kos.axes[1],
      yend = kos.axes[2]
    ),
    arrow = arrow(length = unit(0.02, "npc")),
    color = "royalblue4", size = seg.size
  ) +
  geom_point(
    aes_string(color = row.names(dep.pcd.kos.arws)[1]),
    shape = 19, size = outer.pt.sz
  ) +
  geom_point(
    aes_string(fill = row.names(dep.pcd.kos.arws)[2]),
    shape = 21, size = inner.pt.sz, color = "white", show.legend = F
  ) +
  geom_text_repel(
    data = dep.pcd.kos.arws,
    aes_string(
      x = kos.axes[1],
      y = kos.axes[2],
      label = "var"
    ),
    size = lab.size,
    color = "royalblue4"
  ) +
  scale_color_gradientn(
    name = gsub(" ", "\n", dep.pcd.kos.arws$var[1]), 
    colors = purple_to_green
    ) +
  scale_fill_gradient(
    name = gsub(" ", "\n", dep.pcd.kos.arws$var[2]), 
    low = "white", 
    high = "black"
    ) +
  labs(title = "C", subtitle = "Functional diversity") +
  theme(legend.position = "bottom", axis.title.y = element_blank(), legend.box.just = 1)
leg3 <- get_legend(dep.pcd.kos0)
dep.pcd.kos1 <- dep.pcd.kos0 + theme(legend.position = "none")
dep.pcd.kos2 <- dep.pcd.kos0 + 
  theme(legend.position = "none", axis.title.y = element_text(angle = 90))

plot.wds1 <- c(0.35, 0.325, 0.325)
sum(plot.wds1)
plot.grob1 <- plot_grid(
  imp.kos, ic.pcd.kos, dep.pcd.kos1,
  nrow = 1,
  rel_widths = plot.wds1
)
leg.wds1 <- c(0.25, 0.50, 0.25)
sum(leg.wds1)
leg.grob1 <- plot_grid(leg1, leg2.1, leg3, nrow = 1, rel_widths = leg.wds1)

figure1 <- plot_grid(plot.grob1, leg.grob1, ncol = 1, rel_heights = c(0.9, 0.1))
print(figure1)
ggsave(
  figure1,
  file = file.path(figDir, "mainFig_chb_par_intrxns_ccas_v1.pdf"),
  width = pnas.2col, height = pnas.max.ht / 2
)

leg.hts2 <- c(0.25, 0.50, 0.25)
sum(leg.hts2)
leg.grob2 <- plot_grid(
  NULL, leg1, NULL, leg2.2, NULL, leg3, NULL,
  ncol = 1,
  # rel_heights = leg.hts2,
  axis = "trbl",
  align = "v"
  )

figure2 <- plot_grid(
  plot_grid(imp.kos, ic.pcd.kos, nrow = 1), 
  plot_grid(dep.pcd.kos2, NULL, leg.grob2, nrow = 1, rel_widths = c(0.5, 0.1, 0.4)), 
  ncol = 1
  )
print(figure2)
``` 

## 5. Compound Poisson Generalized Linear Models

See code in Rmd. These chunks are best run on a machine with many cores and a good amount of storage.

```{r cpglm-condsInModel-gen, eval=FALSE}
# This analysis is better run on a machine with many cores to take advantage of parallelization
{set.seed(main.seed)

terms.file <- file.path(saveDir, "list_analysis_terms_envfit.rds")
nCores <- 2 # ideally set to more, especially for kos
type <- "tax" # "kos" "tax"
dfsDir <- paste("CondsInModel", type, "DFs", sep = "_")
if (!dir.exists(file.path(saveDir, dfsDir))) {
  dir.create(file.path(saveDir, dfsDir))
}
# print(type, quote=F)
phy.name <- paste("stool1", type, "phy", sep = ".")
r2.data <- readRDS(file.path(saveDir, "df_envfit_axes12_r2_data.rds"))
child.r2s <- subset(r2.data, grepl("^child_behav_", var) & t == type)
child.ordered <- child.r2s$var[order(child.r2s$r2, decreasing = TRUE)]
esa.r2s <- subset(r2.data, grepl("^esa_", var) & t == type)
esa.ordered <- esa.r2s$var[order(esa.r2s$r2, decreasing = TRUE)]
analysis.terms <- readRDS(terms.file)[[type]]
vars <- c(child.ordered, esa.ordered)
conds <- c(
  analysis.terms[["demo_"]], 
  analysis.terms[["gut_"]],
  analysis.terms[["Diet_"]]
  )
conds <- conds[!is.na(conds)]

phy.obj <- eval(parse(text = phy.name))
pruned.phy <- prune_samples(
  row.names(sample_data(phy.obj)[complete.cases(sample_data(phy.obj)[, vars]), ]),
  phy.obj
)
pruned.phy <- prune_taxa(taxa_sums(pruned.phy) > 0, pruned.phy)
metadata <- as(sample_data(pruned.phy)[, c(vars, conds)], "data.frame")
comm.tbl <- as(otu_table(pruned.phy), "matrix")

if (type == "tax") {
  colnames(comm.tbl) <- gsub("|", ".", colnames(comm.tbl), fixed = T)
}

cl <- makeCluster(nCores, type = "FORK")
registerDoParallel(cl, nCores)
all.units <- colnames(comm.tbl)

# bad.pairs <- data.frame(
#   "var" = c(
#     "esa_LEC_Turmoil_Combined",
#     "esa_LEC_TOT_Combined",
#     "child_behav_CBQ_Sadness_t1"
#   ),
#   "unit" = c(
#     "k__Bacteria.p__Bacteroidetes.c__Bacteroidia.o__Bacteroidales.f__Bacteroidaceae.g__Bacteroides.s__Bacteroides_intestinalis.t__GCF_000172175",
#     "k__Bacteria.p__Bacteroidetes.c__Bacteroidia.o__Bacteroidales.f__Bacteroidaceae.g__Bacteroides.s__Bacteroides_intestinalis.t__GCF_000172175",
#     "k__Bacteria.p__Firmicutes.c__Bacilli.o__Lactobacillales.f__Streptococcaceae.g__Streptococcus.s__Streptococcus_australis.t__Streptococcus_australis_unclassified"
#   )
# )

### For testing
# var <- "esa_LEC_Turmoil_Combined"
# unit <- "k__Bacteria.p__Bacteroidetes.c__Bacteroidia.o__Bacteroidales.f__Bacteroidaceae.g__Bacteroides.s__Bacteroides_intestinalis.t__GCF_000172175"
# unit <- all.units[1]
# var <- vars[1]

all.pvals <- foreach(
  unit = all.units,
  .final = function(x) setNames(x, all.units),
  .verbose = TRUE
) %dopar% {
  sapply(
    vars,
    function(var) {
      model.data <- cbind(metadata[, c(var, conds), drop = F], comm.tbl[, unit, drop = F])
      unit.frm <- paste(unit, "~", paste(c(var, conds), collapse = " + "))
      # if (var %in% bad.pairs$var & unit %in% bad.pairs$unit) {
      #   unit.model <- NA
      #   class(unit.model) <- "try-error"
      # } else {
        unit.model <- try(
          cpglm(
            as.formula(unit.frm), 
            data = model.data, 
            link = "identity", 
            optimizer = "L-BFGS-B"
            ), 
          silent = T
          )
      # }
      if (class(unit.model) == "try-error") {
        unit.pval <- NA
        intercept <- NA
        slope <- NA
      } else {
        sink(file = "tmp.txt")
        model.sum <- summary(unit.model)
        sink()
        unit.pval <- model.sum$coefficients[var, "Pr(>|t|)"]
        intercept <- unname(unit.model$coefficients[1])
        slope <- unname(unit.model$coefficients[var])
      }
      ### ^ There is no r2 value for this method (AIC-based), so going to use slope instead
      unit.df <- data.frame(
        "unit" = unit,
        "b" = intercept,
        "m" = slope
      )
      saveRDS(
        unit.df, 
        file.path(saveDir, dfsDir, paste("df_", var, "-", unit, ".rds", sep = ""))
        )
      return(unit.pval)
    }
  )
}
stopCluster(cl)
}
saveRDS(
  all.pvals,
  file = file.path(
    saveDir,
    paste("list_", type, "_childSesVars_allUnits_cpglms_condsInModel_unadjPvals.rds",
          sep = ""
    )
  )
)
```

<!-- ### Taxon-Covariate Relationships -->

```{r cpglm-condsInModel-tax-plots-gen, eval=FALSE}
dfsDir <- "CondsInModel_tax_DFs"
dfsTar <- paste(dfsDir, "tgz", sep = ".")
varPlotDir <- file.path(saveDir, "cpglm_condsInModel_tax_plots")
if (!dir.exists(varPlotDir)) {
  dir.create(varPlotDir)
}
# unitPlotDir <- file.path(varPlotDir, "unit_plots")
# if(!dir.exists(unitPlotDir)) {
#     dir.create(unitPlotDir)
# }
tax.unadj.pvals <- readRDS(
  file = file.path(saveDir, "list_tax_childSesVars_allUnits_cpglms_condsInModel_unadjPvals.rds")
)

# tax.unadj.pvals %>% head
tax.unadj.pvals.df <- data.frame(tax.unadj.pvals, stringsAsFactors = F) %>%
  as.matrix() %>%
  melt(as.is = T)
tax.adj.pvals <- data.frame(
  "var" = tax.unadj.pvals.df$Var1,
  "unit" = tax.unadj.pvals.df$Var2,
  "adj.pval" = p.adjust(tax.unadj.pvals.df$value, method = "fdr"),
  stringsAsFactors = F
)
tax.sig.pvals <- droplevels(subset(tax.adj.pvals, adj.pval < 0.05))
if (nrow(tax.sig.pvals) == 0) {
  print("No Taxa have a significant relationship with any co-variates")
} else {
  # metadata <- readRDS(file.path(saveDir, "df_allVars_condsResids.rds"))
  metadata <- as(sample_data(stool1.tax.phy), "data.frame")
  comm.tbl <- as(otu_table(stool1.tax.phy), "matrix")
  comm.tbl <- comm.tbl[row.names(metadata), ]
  colnames(comm.tbl) <- gsub("|", ".", colnames(comm.tbl), fixed = TRUE)
  tax.sig.pvals$spearman.r <- apply(
    tax.sig.pvals[, 1:2], 1,
    function(row) {
      # print(row)
      cor(as.numeric(metadata[, as.character(row["var"])]),
          comm.tbl[, as.character(row["unit"])],
          method = "spearman",
          use = "complete.obs"
      )
    }
  )
  tax.plot.list <- list()
  unit.label.list <- list()
  # v <- "child_behav_CBCL_extern_t1"
  for (v in sort(unique(tax.sig.pvals$var))) {
    print(v)
    var.sig.pvals <- subset(tax.sig.pvals, var == v) %>% droplevels()
    if (nrow(var.sig.pvals) == 0) {
      NULL
    } else {
      var.data <- cbind(
        metadata[, v, drop = F],
        comm.tbl[, as.character(var.sig.pvals$unit), drop = F]
      ) %>%
        melt(id.var = v, variable.name = "unit", value.name = "rel.abund")
      var.data$unit.short <- strsplit(as.character(var.data$unit), ".", fixed = T) %>%
        sapply(tail, 2) %>%
        apply(2, paste, collapse = "\n")
      unit.med.relAbund <- doBy::summaryBy(rel.abund ~ unit, data = var.data, FUN = median)
      good.units <- subset(unit.med.relAbund, rel.abund.median > 0)
      if (nrow(good.units) == 0) {
        NULL
      } else {
        var.data <- subset(var.data, unit %in% good.units$unit) %>% droplevels()
        cor.data <- subset(var.sig.pvals, var == v & unit %in% unique(var.data$unit)) %>%
          droplevels()
        unit.label.list[[v]] <- data.frame(
          "var" = v,
          "relationship" = ifelse(cor.data$spearman.r < 0, "-",
                                  ifelse(cor.data$spearman.r > 0, "+", "flat")
          ),
          "taxon" = unique(var.data$unit)
        )
        reg.lines <- data.frame(NULL)
        # u <- unique(as.character(var.data$unit))[1]
        reg.slopes <- data.frame(
          matrix(ncol = 2, nrow = length(unique(as.character(var.data$unit))))
        ) %>% setNames(c("unit", "reg.slope"))
        row.names(reg.slopes) <- unique(as.character(var.data$unit))
        for (u in unique(as.character(var.data$unit))) {
          print(u)
          model.data <- subset(var.data, unit == u)
          file.vu <- paste(dfsDir, "/df_", v, "-", u, ".rds", sep = "")
          untar(file.path(saveDir, dfsTar),
                files = file.vu,
                compressed = "gzip"
          )
          unit.df <- readRDS(file.vu)
          file.remove(file.vu)
          file.remove(dfsDir)
          unit.df$unit.short <- strsplit(as.character(u), ".", fixed = T) %>%
            sapply(tail, 2) %>%
            apply(2, paste, collapse = "\n")
          reg.slopes[u, c("unit", "reg.slope")] <- unit.df[1, c("unit.short", "m")]
          reg.lines <- rbind(reg.lines, unit.df)
        }
        unit.label.list[[v]]$reg.slope <- reg.slopes$reg.slope
        var.data$log.rel.abund <- log(var.data$rel.abund)
        var.plot <- ggplot(data = var.data, aes_string(x = v, y = "log.rel.abund")) +
          geom_abline(
            data = reg.lines, aes(intercept = b, slope = m),
            color = "blue", size = 0.5
          ) +
          geom_point(aes(shape = ifelse(rel.abund == 0, "zero", "not zero")),
                     alpha = 0.6
          ) +
          scale_shape_manual(name = "Relative abundance", values = c(16, 1)) +
          labs(
            y = "log(Unit Relative Abundance)",
            x = mgsub(vars.legend$var.name, vars.legend$description, v),
            title = "Compound Poisson GLMs on variable residuals"
          ) +
          facet_wrap(~unit.short, scales = "free_y") +
          theme(
            axis.text.y = element_blank(),
            legend.position = "top",
            legend.title = element_text(size = 8),
            legend.text = element_text(size = 6),
            axis.text = element_text(size = 8),
            axis.title = element_text(size = 10),
            strip.text = element_text(size = 8)
          )
        tax.plot.list[[v]] <- var.plot
        nCol <- ceiling(sqrt(length(unique(var.data$unit))))
        nRow <- ceiling(length(unique(var.data$unit)) / nCol)
        ggsave(
          var.plot,
          file = file.path(varPlotDir, paste(gsub(".", "_", v, fixed = T), "png", sep = ".")),
          height = nRow * 2,
          width = nCol * 2.5,
          dpi = 300
        )
      }
    }
  }
  
  # unit.label.df <- do.call("rbind", e$unit.label.list)
  unit.label.df <- do.call("rbind", unit.label.list)
  row.names(unit.label.df) <- NULL
  unit.label.df$var <- mgsub(
    c(vars.legend$var.name, ".resid"),
    c(vars.legend$description, " Residuals"),
    unit.label.df$var
  )
  taxa.split <- strsplit(as.character(unit.label.df$taxon), ".", fixed = TRUE)
  taxa.tbl <- sapply(taxa.split, function(i) {
    if (length(i) < 8) {
      i <- c(i, "t_NA")
    }
    i
  }) %>%
    t() %>%
    as.data.frame()
  taxa.tbl <- taxa.tbl[, -8]
  names(taxa.tbl) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
  names(unit.label.df) <- c("Variable", "Relationship", "lab", "Slope")
  unit.label.df <- cbind(unit.label.df[, c(1, 2, 4)], taxa.tbl)
  unit.label.df[, -c(1:3)] <- apply(
    unit.label.df[, -c(1:3)], 2,
    function(col) sapply(strsplit(col, "__"), `[`, 2)
  )
  # write.table(
  #   unit.label.df,
  #   file = file.path(saveDir, "df_sigVarTaxRels_condsInModel.txt"),
  #   sep = "\t", quote = F, row.names = F
  # )
  saveRDS(unit.label.df, file = file.path(saveDir, "df_sigVarTaxRels_condsInModel.rds"))
  unit.label.dt <- as.data.table(unit.label.df)
  saveRDS(unit.label.dt, file = file.path(saveDir, "dt_sigVarTAXrels_condsInModel.rds"))
  tax.plot.list <- tax.plot.list[!sapply(tax.plot.list, is.null)]
  nRows <- sapply(tax.plot.list, function(plot) {
    panels <- max(as.numeric(ggplot_build(plot)$data[[1]]$PANEL))
    if (panels > 3) {
      nCol <- ceiling(sqrt(panels))
      ceiling(panels / nCol)
    } else {
      1
    }
  })
  plot.heights <- nRows / sum(nRows)
  tax.plot.height <- ifelse(sum(nRows) / 4 < length(nRows),
                            length(nRows) * 6,
                            (sum(nRows) / 4) * 8
  )
}
```

```{r tax-network-cap}
tax.network.cap <- ""
```

```{r cpglm-tax-network, fig.width=pnas.2col/0.9050, fig.height=pnas.max.ht, out.height=650*1.2, fig.cap=tax.network.cap, eval=FALSE}
tax.vars.rels <- readRDS(file.path(saveDir, "dt_sigVarTAXrels_condsInModel.rds"))

plot.data0 <- tax.vars.rels[
  , 
  .(taxComps = length(Species), Species, Slope), 
  by = Variable
  ]
plot.data <- plot.data0[
  , 
  .(varComps = length(Variable), taxComps, Variable, Slope), 
  by = Species
  ]
plot.data$Type <- ifelse(grepl("CBCL|CBQ", plot.data$Variable), "Child Behavior", "ESA")
# plot.data$Species <- gsub("_", " ", plot.data$Species)

ordered.taxa <- levels(with(plot.data, reorder(Species, varComps)))
ordered.chb <- levels(with(plot.data[Type != "ESA"], reorder(Variable, taxComps)))
ordered.esa <- levels(with(plot.data[Type == "ESA"], reorder(Variable, taxComps)))
range <- c(-5, 5)
l <- mean(range) - 1.5
r <- mean(range) + 1.5
seg.data <- within(plot.data, {
  x0 <- ifelse(Type != "ESA", range[1], range[2])
  y0 <- ifelse(Type != "ESA",
               (match(Variable, ordered.chb) - 0.5) / length(ordered.chb),
               (match(Variable, ordered.esa) - 0.5) / length(ordered.esa)
  )
  x1 <- ifelse(Type != "ESA", l, r)
  y1 <- (match(Species, ordered.taxa) - 0.5) / length(ordered.taxa)
  mid <- mean(range)
})
seg.data <- seg.data[order(abs(Slope))]
lab.font.pt <- 6
lab.size <- lab.font.pt * 0.35
taxa.line <- ggplot(data = seg.data) +
  geom_segment(aes(x = x0, xend = x1, y = y0, yend = y1, color = Slope, size = abs(Slope), alpha = abs(Slope))) +
  scale_color_gradientn(name = "Reg. Slope", colors = blue_to_red) +
  scale_size_continuous(name = "abs(Reg. Slope)", range = c(0.05, 1.5)) +
  scale_alpha_continuous(name = "abs(Reg. Slope)", range = c(0.6, 0.8)) +
  geom_tile(aes(x = mid, y = y1), width = r - l, fill = "white", color = "black") +
  geom_text(
    data = unique(seg.data[, .(Species, mid, y1)]), aes(x = mid, y = y1, label = Species),
    size = lab.size, fontface = "italic"
  ) +
  geom_tile(
    data = seg.data[Type != "ESA"], aes(x = x0 + l, y = y0),
    width = r - l, height = 0.05, fill = "white", color = "black"
  ) +
  geom_text(
    data = unique(seg.data[Type != "ESA", .(Variable, x0, y0)]),
    aes(x = x0 + l, y = y0, label = Variable), size = lab.size
  ) +
  geom_tile(
    data = seg.data[Type == "ESA"], aes(x = x0 + r, y = y0),
    width = r - l, height = 0.05, fill = "white", color = "black"
  ) +
  geom_text(
    data = unique(seg.data[Type == "ESA", .(Variable, x0, y0)]),
    aes(x = x0 + r, y = y0, label = Variable), size = lab.size
  ) +
  scale_x_continuous(
    breaks = c(min(seg.data$x0) + l, 0, max(seg.data$x0) + r),
    labels = c("Behavioral Dysregulation", "Taxa", "Socioeconomic Risk"),
    position = "top"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    panel.background = element_blank(),
    panel.border = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    axis.line = element_blank(),
    legend.position = "bottom",
    legend.key.size = unit(11, "pt"),
    legend.box.margin = unit(c(0, 0, 0, 0), "pt"),
    legend.box.spacing = unit(5, "pt")
  )
taxa.line
fig.wd <- pnas.2col / 0.8887
ggsave(taxa.line, file = file.path(figDir, "mainFig_chb_esa_tax_regNetwork.pdf"), width = fig.wd, height = 6)
system("open -a /Applications/Inkscape.app/ Figure_plots/mainFig_chb_esa_tax_regNetwork.pdf")
```

<!-- ### KO-Covariate Relationships -->

```{r cpglm-condsInModel-kos-plots-gen, eval=FALSE}
dfsDir <- "CondsInModel_kos_DFs"
dfsTar <- paste(dfsDir, "tgz", sep = ".")
varPlotDir <- file.path(saveDir, "cpglm_condsInModel_kos_plots")
if (!dir.exists(varPlotDir)) {
  dir.create(varPlotDir)
}
unitPlotDir <- file.path(varPlotDir, "unit_plots")
if (!dir.exists(unitPlotDir)) {
  dir.create(unitPlotDir)
}
kos.unadj.pvals <- readRDS(
  file = file.path(
    saveDir, "list_kos_childSesVars_allUnits_cpglms_condsInModel_unadjPvals.rds"
  )
)
kos.unadj.pvals.df <- data.frame(kos.unadj.pvals, stringsAsFactors = F) %>%
  as.matrix() %>%
  melt(as.is = T)
kos.adj.pvals <- data.frame(
  "var" = kos.unadj.pvals.df$Var1,
  "unit" = kos.unadj.pvals.df$Var2,
  "adj.pval" = p.adjust(kos.unadj.pvals.df$value, method = "fdr"),
  stringsAsFactors = F
)
kos.sig.pvals <- droplevels(subset(kos.adj.pvals, adj.pval < 0.05))
if (nrow(kos.sig.pvals) == 0) {
  print("No KOs have a significant relationship with any co-variates")
} else {
  metadata <- as(sample_data(stool1.kos.phy), "data.frame")
  comm.tbl <- as(otu_table(stool1.kos.phy), "matrix")
  comm.tbl <- comm.tbl[row.names(metadata), ]
  kos.sig.pvals$spearman.r <- apply(
    kos.sig.pvals[, 1:2], 1,
    function(row) {
      cor(as.numeric(metadata[, row["var"]]),
          comm.tbl[, row["unit"]],
          method = "spearman",
          use = "complete.obs"
      )
    }
  )
  my.kos.mods.clpsd <- readRDS(kegg.map.files[2])
  my.kos.mods.clpsd$modules[my.kos.mods.clpsd$modules == "NA"] <- NA
  my.kos.mods.clpsd$modules.names[my.kos.mods.clpsd$modules.names == "NA"] <- NA
  # write.table(my.kos.mods.clpsd, file="~/Desktop/kos_mods_tbl.txt", sep="\t", quote=F, row.names=F)
  my.kos.mods.clpsd$modules.names[!is.na(my.kos.mods.clpsd$modules.names)] <- paste(
    "MODS:", my.kos.mods.clpsd$modules.names[!is.na(my.kos.mods.clpsd$modules.names)]
  )
  my.kos.mods.clpsd$kos.name[!is.na(my.kos.mods.clpsd$kos.name)] <- paste(
    "KO:", my.kos.mods.clpsd$kos.name[!is.na(my.kos.mods.clpsd$kos.name)]
  )
  kos.sig.pvals.modIDs <- merge(kos.sig.pvals,
                                my.kos.mods.clpsd,
                                by.x = "unit",
                                by.y = "kos",
                                all.x = TRUE
  )
  kos.pfam.clpsd <- readRDS(file = "Static_data/kos_pfam_top5matches_clpsd.rds")
  kos.pfam.clpsd$pfams[!is.na(kos.pfam.clpsd$pfams)] <- paste(
    "PFAMS:", kos.pfam.clpsd$pfams[!is.na(kos.pfam.clpsd$pfams)]
  )
  kos.sig.pvals.modIDs.pfams <- merge(kos.sig.pvals.modIDs,
                                      kos.pfam.clpsd,
                                      by.x = "unit",
                                      by.y = "kos",
                                      all.x = TRUE
  )
  kos.plot.list <- list()
  unit.label.list <- list()
  # v <- "child_behav_CBCL_depression_t1"
  for (v in sort(unique(kos.sig.pvals$var))) {
    print(v)
    var.sig.pvals <- subset(kos.sig.pvals.modIDs.pfams, var == v) %>% droplevels()
    if (nrow(var.sig.pvals) == 0) {
      NULL
    } else {
      var.data <- cbind(
        metadata[, v, drop = F],
        comm.tbl[, as.character(var.sig.pvals$unit), drop = F]
      ) %>%
        melt(id.var = v, variable.name = "unit", value.name = "rel.abund")
      plot.data <- merge(var.data, var.sig.pvals, by = "unit", all.x = T)
      plot.data$log.rel.abund <- log(plot.data$rel.abund)
      unit.med.relAbund <- doBy::summaryBy(rel.abund ~ unit,
                                           data = plot.data,
                                           FUN = median
      )
      good.units <- subset(unit.med.relAbund, rel.abund.median > 0)
      if (nrow(good.units) == 0) {
        NULL
      } else {
        plot.data <- subset(plot.data, unit %in% good.units$unit)
        label.base <- plot.data[, c(
          "unit",
          "spearman.r",
          "kos.name",
          "modules.names",
          "pfams"
        )] %>% unique()
        labs <- NULL
        for (row in c(1:nrow(label.base))) {
          row.lab <- ifelse(
            !is.na(label.base$modules.names[row]),
            label.base$modules.names[row],
            ifelse(
              !is.na(label.base$kos.name[row]),
              paste(label.base$kos.name[row], label.base$pfams[row], sep = " :: "),
              label.base$pfams[row]
            )
          )
          labs <- c(labs, row.lab)
        }
        label.base$plot.label <- labs
        label.data <- merge(
          label.base,
          doBy::summaryBy(log.rel.abund ~ unit, plot.data, FUN = max),
          by = "unit"
        )
        unit.label.list[[v]] <- data.frame(
          "var" = v,
          "unit" = label.data$unit,
          "relationship" = ifelse(label.data$spearman.r < 0, "-",
                                  ifelse(label.data$spearman.r > 0, "+", "flat")
          ),
          "label" = label.data$plot.label
        )
        reg.lines <- data.frame(NULL)
        u <- unique(as.character(plot.data$unit))[1]
        reg.slopes <- data.frame(
          matrix(ncol = 2, nrow = length(unique(as.character(plot.data$unit))))
        ) %>% setNames(c("unit", "reg.slope"))
        row.names(reg.slopes) <- unique(as.character(plot.data$unit))
        for (u in unique(as.character(plot.data$unit))) {
          print(u)
          model.data <- subset(plot.data, unit == u)[, c(2, 3, 11)]
          # unit.df <- readRDS(
          #     file.path(saveDir, dfsDir, paste("df_", v, "-", u, ".rds", sep=""))
          # )
          file.vu <- paste(dfsDir, "/df_", v, "-", u, ".rds", sep = "")
          untar(file.path(saveDir, dfsTar),
                files = file.vu,
                compressed = "gzip"
          )
          unit.df <- readRDS(file.vu)
          file.remove(file.vu)
          file.remove(dfsDir)
          unit.df$unit <- as.character(unit.df$unit)
          reg.slopes[u, c("unit", "reg.slope")] <- unit.df[1, c("unit", "m")]
          reg.lines <- rbind(reg.lines, unit.df)
          unit.plot <- ggplot(data = model.data, aes_string(x = v, y = "log.rel.abund")) +
            geom_abline(intercept = unit.df$b, slope = unit.df$m, color = "blue", size = 0.5) +
            geom_point(aes(shape = ifelse(rel.abund == 0, "zero", "not zero")),
                       alpha = 0.6
            ) +
            geom_text(
              data = subset(label.data, unit == u),
              aes(
                label = ifelse(is.na(plot.label),
                               "",
                               mgsub(c(", ", "; "),
                                     c("\n", "\n"),
                                     plot.label,
                                     fixed = T
                               )
                ),
                x = ifelse(spearman.r < 0,
                           max(plot.data[, v], na.rm = T),
                           min(plot.data[, v], na.rm = T)
                ),
                y = log.rel.abund.max,
                hjust = ifelse(spearman.r < 0, 1, 0)
              ),
              vjust = 1, size = 2.5, color = "red"
            ) +
            scale_shape_manual(name = "Relative abundance", values = c(16, 1)) +
            labs(
              y = paste("log(", u, " Rel. Abund.)", sep = ""),
              x = mgsub(
                c(vars.legend$var.name),
                c(vars.legend$description),
                v
              ),
              title = "Linear models on KO abundance (CPGLM)"
            ) +
            theme(
              legend.position = "top",
              legend.title = element_text(size = 18),
              legend.text = element_text(size = 10),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 16),
              strip.text = element_text(size = 16)
            )
          ggsave(unit.plot,
                 file = file.path(
                   unitPlotDir, paste(gsub(".", "_", v, fixed = T), "-", u, ".png", sep = "")
                 ),
                 height = 6,
                 width = 8,
                 dpi = 300
          )
        }
        unit.label.list[[v]]$reg.slope <- reg.slopes$reg.slope
        var.plot <- ggplot(data = plot.data, aes_string(x = v, y = "log.rel.abund")) +
          geom_abline(
            data = reg.lines, aes(intercept = b, slope = m),
            color = "blue", size = 0.5
          ) +
          geom_point(aes(shape = ifelse(rel.abund == 0, "zero", "not zero")),
                     alpha = 0.6
          ) +
          geom_text(
            data = label.data,
            aes(
              label = ifelse(is.na(plot.label),
                             "",
                             mgsub(c(", ", "; "),
                                   c("\n", "\n"),
                                   plot.label,
                                   fixed = T
                             )
              ),
              x = ifelse(spearman.r < 0,
                         max(plot.data[, v], na.rm = T),
                         min(plot.data[, v], na.rm = T)
              ),
              y = log.rel.abund.max,
              hjust = ifelse(spearman.r < 0, 1, 0)
            ),
            vjust = 1, size = 2.5, color = "red"
          ) +
          scale_shape_manual(name = "Relative abundance", values = c(16, 1)) +
          labs(
            y = "log(Unit Relative Abundance)",
            x = mgsub(
              c(vars.legend$var.name, ".resid"),
              c(vars.legend$description, " Residuals"),
              v
            ),
            title = "Linear models on KO abundance (CPGLM)"
          ) +
          facet_wrap(~unit, scales = "free_y") +
          theme(
            legend.position = "top",
            axis.text.y = element_blank()
          )
        kos.plot.list[[v]] <- var.plot
        nCol <- ceiling(sqrt(length(unique(plot.data$unit))))
        nRow <- ceiling(length(unique(plot.data$unit)) / nCol)
        ggsave(
          var.plot,
          file = file.path(varPlotDir, paste(gsub(".", "_", v, fixed = T), "png", sep = ".")),
          height = nRow * 2,
          width = 8,
          dpi = 300
        )
      }
    }
  }
  # kos.plot.list <- kos.plot.list[!sapply(kos.plot.list, is.null)]
  # unit.label.df <- do.call("rbind", e$unit.label.list)
  unit.label.df <- do.call("rbind", unit.label.list)
  unit.label.df[, 1:4] <- apply(unit.label.df[, 1:4], 2, as.character)
  row.names(unit.label.df) <- NULL
  unit.label.df$var <- mgsub(
    c(vars.legend$var.name, ".resid"),
    c(vars.legend$description, " Residuals"),
    unit.label.df$var
  )
  names(unit.label.df) <- c("Variable", "Unit", "Relationship", "Label", "Slope")
  write.table(
    unit.label.df,
    file = file.path(saveDir, "df_sigVarKOrels_condsInModel.txt"),
    sep = "\t", quote = F, row.names = F
  )
  saveRDS(unit.label.df, file = file.path(saveDir, "df_sigVarKOrels_condsInModel.rds"))
  unit.label.dt <- as.data.table(unit.label.df)
  mods.label.dt <- unit.label.dt[grepl("MODS", Label, ignore.case = F)]
  write.table(mods.label.dt,
              file = file.path(saveDir, "df_sigVarMODrels_condsInModel.txt"),
              sep = "\t", quote = F, row.names = F
  )
  saveRDS(mods.label.dt, file = file.path(saveDir, "dt_sigVarMODrels_condsInModel.rds"))
  
  nRows <- sapply(kos.plot.list, function(plot) {
    panels <- max(as.numeric(ggplot_build(plot)$data[[1]]$PANEL))
    nCol <- ceiling(sqrt(panels))
    ceiling(panels / nCol)
  })
  plot.heights <- nRows / sum(nRows)
  kos.plot.height <- ifelse(sum(nRows) / 4 < length(nRows),
                            length(nRows) * 8,
                            (sum(nRows) / 4) * 6
  )
}
```

```{r kos-network-cap}
kos.network.cap <- ""
```

```{r cpglm-kos-network, fig.width=pnas.2col/0.9050, fig.height=pnas.max.ht/0.9230, out.height=650*1.2, fig.cap=kos.network.cap, eval=FALSE}
fig.wd <- pnas.2col / 0.8887
fig.ht <- pnas.max.ht / 0.965
lab.font.pt <- 6
mods.vars.rels <- readRDS(file.path(saveDir, "dt_sigVarMODrels_condsInModel.rds"))
### remove animal/plant-only Mods
mods.vars.rels <- subset(mods.vars.rels, !(Unit %in% c("K06225", "K03064")))
mods.vars.rels$Label <- sapply(
  strsplit(sub("MODS: ", "", unlist(strsplit(mods.vars.rels$Label, " KO:"))), ","),
  `[`, 1
)
#
plot.data0 <- mods.vars.rels[, .(modComps = length(Label), Label, Slope), by = Variable]
plot.data <- plot.data0[
  , 
  .(varComps = length(Variable), modComps, Variable, Slope), 
  by = Label
  ]
plot.data$Type <- ifelse(grepl("CBCL|CBQ", plot.data$Variable), "Child Behavior", "ESA")
# plot.data$Label <- strtrim(plot.data$Label, median(nchar(plot.data$Label)))
# grep("manno-heptose", unique(plot.data$Label), value=F)

new.par <- par(ps = lab.font.pt, family = "Helvetica", fin = c(fig.wd, fig.ht))
to.replace <- data.frame(
  "word" = c(
    "pathway", "; ", "beta", "biosynthesis",
    "system", "polysaccharide", "regulatory", "transport",
    "two", "component", "degradation", "pathogenicity",
    "signature", "deoxyribonuleotide", "zinc", "manganese",
    "; Cell cycle - G2/M transition", "antimicrobial peptide", "sulfate", "Putative",
    "phosphate", "hydroxymethylpyrimidine", "oligosaccharyltransferase", "Oxidation",
    "semialdehyde", "Entner-Doudoroff", "ADP-L-glycero-D-manno-heptose",
    "BRCA1-associated genome surveillance complex ", "; MRX complex"
  ),
  "replace" = c(
    "ptwy", ";", "", "biosyn",
    "sys", "polysac", "reg", "transp",
    "2", "comp", "degrad", "pathogenic",
    "signtr", "dNMP", "Zn", "Mn",
    "", "AMP", "SO4", "Put.",
    "PO4", "HMP", "OST", "oxid'n",
    "semialdhd", "ED", "ADP-LDHep",
    "", ""
  ),
  stringsAsFactors = F
)
max.wd <- 1.26
for (l in 1:length(plot.data$Label)) {
  lab <- plot.data$Label[l]
  lab <- sub("\\(.+\\)[ ;] *", "", lab)
  wd0 <- strwidth(lab, units = "inches")
  if (wd0 > max.wd) {
    lab <- mgsub(to.replace$word, to.replace$replace, lab)
    wd1 <- strwidth(lab, units = "inches")
    if (wd1 > max.wd) {
      lab <- sub(";.+", "", lab)
      wd2 <- strwidth(lab, units = "inches")
      if (wd2 > max.wd) {
        print(lab, quote = F)
      }
    }
  }
  plot.data$Label[l] <- lab
}
par(.pardefault)
ordered.mods <- levels(with(plot.data, reorder(Label, varComps)))
ordered.chb <- levels(with(plot.data[Type != "ESA"], reorder(Variable, modComps)))
ordered.esa <- levels(with(plot.data[Type == "ESA"], reorder(Variable, modComps)))
range <- c(-5, 5)
l <- mean(range) - 1.5
r <- mean(range) + 1.5
seg.data <- within(plot.data, {
  x0 <- ifelse(Type != "ESA", range[1], range[2])
  y0 <- ifelse(Type != "ESA",
               (match(Variable, ordered.chb) - 0.5) / length(ordered.chb),
               (match(Variable, ordered.esa) - 0.5) / length(ordered.esa)
  )
  x1 <- ifelse(Type != "ESA", l, r)
  y1 <- (match(Label, ordered.mods) - 0.5) / length(ordered.mods)
  mid <- mean(range)
})
seg.data$Variable <- sub("/", "/\n", seg.data$Variable)
seg.data <- seg.data[order(abs(Slope))]
lab.size <- lab.font.pt * 0.35
mods.line <- ggplot(data = seg.data) +
  geom_segment(
    aes(
      x = x0, xend = x1, 
      y = y0, yend = y1, 
      color = Slope, 
      size = abs(Slope), 
      alpha = abs(Slope)
      )
    ) +
  scale_color_gradientn(name = "Reg. Slope", colors = blue_to_red) +
  scale_size_continuous(name = "abs(Reg. Slope)", range = c(0.05, 1.5)) +
  scale_alpha_continuous(name = "abs(Reg. Slope)", range = c(0.6, 0.8)) +
  geom_tile(aes(x = mid, y = y1), width = r - l, fill = "white", color = "black") +
  geom_text(
    data = unique(seg.data[, .(Label, mid, y1)]),
    aes(x = mid, y = y1, label = Label),
    size = lab.size
  ) +
  geom_tile(
    data = seg.data[Type != "ESA"], aes(x = x0 + l, y = y0),
    width = r - l, height = 0.05, fill = "white", color = "black"
  ) +
  geom_text(
    data = unique(seg.data[Type != "ESA", .(Variable, x0, y0)]),
    aes(x = x0 + l, y = y0, label = Variable),
    size = lab.size
  ) +
  geom_tile(
    data = seg.data[Type == "ESA"], aes(x = x0 + r, y = y0),
    width = r - l, height = 0.05, fill = "white", color = "black"
  ) +
  geom_text(
    data = unique(seg.data[Type == "ESA", .(Variable, x0, y0)]),
    aes(x = x0 + r, y = y0, label = Variable),
    size = lab.size
  ) +
  scale_x_continuous(
    breaks = c(min(seg.data$x0) + l, 0, max(seg.data$x0) + r),
    labels = c("Behavioral Dyregulation", "Functional Modules", "Socioeconomic Risk"),
    position = "top"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    panel.border = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    axis.line = element_blank(),
    legend.position = "bottom",
    legend.key.size = unit(11, "pt"),
    legend.box.margin = unit(c(0, 0, 0, 0), "pt"),
    legend.box.spacing = unit(5, "pt")
  )
mods.line

ggsave(mods.line, file=file.path(figDir, "mainFig_chb_esa_kos_regNetwork.pdf"), width=fig.wd, height=fig.ht)
system("open -a /Applications/Inkscape.app/ Figure_plots/mainFig_chb_esa_kos_regNetwork.pdf")

###  ISSC Talk FIG
# fig.wd <- pnas.2col/0.8887
mods.leg <- cowplot::get_legend(mods.line)
mods.line <- mods.line + theme(legend.position = "none")
ggsave(mods.line, file = file.path(figDir, "mainFig_chb_esa_kos_regNetwork.pdf"), width=fig.wd, height=fig.ht)
ggsave(mods.leg, file = file.path(figDir, "mainFig_chb_esa_kos_regNetwork_legend.pdf"))
system("open -a /Applications/Inkscape.app/ Figure_plots/mainFig_chb_esa_kos_regNetwork.pdf")
# ggsave(mods.line,
#        file = "~/Google Drive/OSU/OSU_Home_Sync/ISSC_talk/mainFig_chb_esa_kos_regNetwork.pdf",
#        width = 190, height = 190 * (fig.ht / fig.wd), units = "mm"
# )
# # system("open '/Users/stagamak/Google Drive/OSU/OSU_Home_Sync/ISSC_talk/mainFig_chb_esa_kos_regNetwork.pdf'")
# system("open -a /Applications/Inkscape.app/ '/Users/stagamak/Google Drive/OSU/OSU_Home_Sync/ISSC_talk/mainFig_chb_esa_kos_regNetwork.pdf'")

```

<!-- ### T6SS KOs/Modules and Taxa Relationships -->

```{r t6ss-taxa, fig.height=pnas.2col, eval=FALSE}
set.seed(main.seed)
kos.df <- readRDS(file = file.path(saveDir, "df_sigVarKOrels_condsInModel.rds"))
kos.dt <- as.data.table(kos.df)
t6ss.dt <- kos.dt[grepl("type VI", Label, ignore.case = T)]
t6ss.kos <- sort(unique(t6ss.dt$Unit))
kos.tbl <- as(otu_table(stool1.kos.phy), "matrix")
t6ss.tbl <- kos.tbl[, t6ss.kos]
t6ss.m.dt <- melt(t6ss.tbl, varnames = c("smpl", "kos"), value.name = "rel.kos.abund") %>% as.data.table()
tax.tbl <- as(otu_table(stool1.tax.phy), "matrix")
tax.m.dt <- melt(tax.tbl, varnames = c("smpl", "taxon"), value.name = "rel.tax.abund") %>% as.data.table()
t6ss.tax.data <- merge(tax.m.dt, t6ss.m.dt, by = "smpl", all = T, allow.cartesian = T)
cpglm.res <- data.frame()
# txn <- colnames(tax.tbl)[1]
# t6.ko <- t6ss.kos[1]
for (txn in colnames(tax.tbl)) {
  for (t6.ko in t6ss.kos) {
    comp.data <- t6ss.tax.data[taxon == txn & kos == t6.ko]
    if (median(comp.data$rel.tax.abund) > 0 & median(comp.data$rel.kos.abund) > 0) {
      comp.data$log.kos.abund <- log(comp.data$rel.kos.abund)
      tax.t6.cpglm <- try(
        cpglm(rel.tax.abund ~ log.kos.abund, data = comp.data),
        silent = T
      )
      if (class(tax.t6.cpglm) != "try-error") {
        sink(file = "tmp.txt")
        model.sum <- summary(tax.t6.cpglm)
        sink()
        pval <- model.sum$coefficients[2, "Pr(>|t|)"]
        slope <- unname(tax.t6.cpglm$coefficients[2])
        intercept <- unname(tax.t6.cpglm$coefficients[1])
        cpglm.df <- data.frame(
          "taxon" = txn,
          "kos" = t6.ko,
          "m" = slope,
          "b" = intercept,
          "unadj.p" = pval
        )
        cpglm.res <- rbind(cpglm.res, cpglm.df)
      }
    }
  }
}
# View(cpglm.res)
cpglm.res$adj.p <- p.adjust(cpglm.res$unadj.p, method = "fdr")
cpglm.res <- as.data.table(cpglm.res)
sig.assoc <- cpglm.res[adj.p <= 0.05]
# View(sig.assoc)
sig.assoc$n <- LETTERS[1:nrow(sig.assoc)]
sig.assoc <- within(sig.assoc, {
  tax.species <- str_extract(taxon, "s__[a-zA-Z0-9 _]+")
  tax.species <- mgsub(c("s__", "_bacterium"), c("", "", "715"), tax.species)
  # plot.lab <- paste(n, " ", tax.species, "\n", kos, sep="")
})
sig.assoc$plot.lab <- factor(sig.assoc$plot.lab, levels = unique(sig.assoc$plot.lab))
# # sig.cors$taxon
sig.t6ss.tax <- t6ss.tax.data[sig.assoc, on = c(taxon = "taxon", kos = "kos")]
sig.t6ss.tax <- within(sig.t6ss.tax, {
  shp <- ifelse(rel.kos.abund == 0 & rel.tax.abund == 0, "zero-zero",
                ifelse(rel.kos.abund == 0, "ko-zero",
                       ifelse(rel.tax.abund == 0, "tax-zero", "no-zero")
                )
  )
})

t6ss.tax.plot <- ggplot(data = sig.t6ss.tax) +
  geom_abline(
    data = sig.assoc, aes(intercept = b, slope = m),
    color = "blue", size = 0.5
  ) +
  geom_point(aes(x = log(rel.kos.abund), y = log(rel.tax.abund), shape = shp), alpha = 0.6) +
  scale_shape_manual(name = "Relative abundance", values = c(16, 1), labels = c("not zero", "zero")) +
  labs(
    y = "log(Taxon Relative Abundance)",
    x = "log(KO Relative Abundance)"
  ) +
  facet_wrap(~n, scales = "free") +
  theme(
    legend.position = "top",
    strip.text = element_text(hjust = 0)
  )
plot.grob <- arrangeGrob(t6ss.tax.plot)
t6ss.dt$Tbl.Label <- mgsub(c(":: ", "; "), c("\n", "\n    "), t6ss.dt$Label)

tax.kos <- sig.assoc[, .(Panel = n, Taxon = gsub("_", " ", tax.species), KO = kos)]
ff <- matrix(c(1, 3, 1), ncol = ncol(tax.kos), nrow = nrow(tax.kos), byrow = T)
tt1 <- ttheme_default(base_size = 8, core = list(fg_params = list(fontface = as.vector(ff))))
tax.ko.tbl <- tableGrob(tax.kos,
                        rows = NULL, theme = tt1
)

kos.labs <- unique(t6ss.dt[order(Unit), .(KO = Unit, Label = Tbl.Label)])
hj <- matrix(c(0.5, 0), ncol = ncol(kos.labs), nrow = nrow(kos.labs), byrow = T)
x <- matrix(c(0.5, 0.01), ncol = ncol(kos.labs), nrow = nrow(kos.labs), byrow = T)

tt2 <- ttheme_default(
  core = list(fg_params = list(hjust = as.vector(hj), x = as.vector(x))),
  colhead = list(fg_params = list(hjust = c(0.5, 0), x = c(0.5, 0.01))),
  base_size = 8
)
ko.lab.tbl <- tableGrob(kos.labs, rows = NULL, theme = tt2)
table.grob <- arrangeGrob(tax.ko.tbl, ko.lab.tbl, nrow = 1)

t6ss.fig <- arrangeGrob(plot.grob, table.grob,
                        nrow = 2,
                        heights = c(0.6, 0.4)
)
ggsave(t6ss.fig,
       file = file.path(figDir, "suppFig_t6ss_tax_regs.pdf"),
       width = pnas.2col + 2.5, height = pnas.max.ht + 2
)
system(paste("open", file.path(figDir, "suppFig_t6ss_tax_regs.pdf")))
system(paste("open -a /Applications/Inkscape.app", file.path(figDir, "suppFig_t6ss_tax_regs.pdf")))
```

```{r addressing-some-reviewer-comments, eval=FALSE}
smpl.data <- as(sample_data(stool1.kos.phy), "data.frame")
ggplot(smpl.data, aes(x = parent_PSI_PC_dysfunc_t1, y = esa_LEC_Turmoil_Combined)) + 
  geom_point()
cor.test(smpl.data$parent_PSI_PC_dysfunc_t1, smpl.data$esa_LEC_Turmoil_Combined)
cor.test(
  smpl.data$parent_PSI_PC_dysfunc_t1, 
  smpl.data$esa_LEC_Turmoil_Combined, 
  method = "spearman"
  )

ggplot(smpl.data, aes(x = parent_PSI_PC_dysfunc_t1, y = child_behav_CBCL_depression_t1)) + 
  geom_point()
ggsave(file = "Figure_plots/parent-child_dys_vs_CBCL-depression_plot.pdf")
cor.test(smpl.data$parent_PSI_PC_dysfunc_t1, smpl.data$child_behav_CBCL_depression_t1)
cor.test(
  smpl.data$parent_PSI_PC_dysfunc_t1, 
  smpl.data$child_behav_CBCL_depression_t1, 
  method = "spearman"
  )
```

